---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
set.seed(1)
if (!requireNamespace("contentvalidR", quietly = TRUE)) {
  devtools::load_all(quiet = TRUE)
}
```

# contentvalidR

<!-- badges: start -->
[![R-CMD-check](https://github.com/JUhalt/contentvalidR/actions/workflows/R-CMD-check.yaml/badge.svg)](https://github.com/JUhalt/contentvalidR/actions/workflows/R-CMD-check.yaml)
[![License: MIT](https://img.shields.io/badge/License-MIT-yellow.svg)](LICENSE)
[![Lifecycle: experimental](https://img.shields.io/badge/lifecycle-experimental-orange.svg)](https://lifecycle.r-lib.org/articles/stages.html)
<!-- badges: end -->

**contentvalidR** provides quantitative tools for substantive and content-oriented scale pretesting. The package is being developed around three complementary workflows:

1. **Item sorting** — Anderson & Gerbing (1991) Psa/Csv, exact item-level inference following Howard & Melloy (2016), and scale-level empirical norms from Colquitt et al. (2019).
2. **Construct ratings** — Hinkin & Tracey (1999) fully crossed ratings, HTC/HTD, Greenhouse-Geisser-aware repeated-measures item screening, and Colquitt et al. (2019) scale-level norms.
3. **Expert panels** — Aiken's V, Lawshe CVR, CVI/modified kappa, and item-objective congruence.

The design goal is **interpretable output rather than coefficient dumps**. Recommended workflow functions summarize what the evidence supports, flag items that need attention, and distinguish statistical screening from substantive decisions.

> Quantitative content-validity statistics are one part of a broader validity argument. They complement, rather than replace, construct definition, domain coverage, qualitative expert feedback, cognitive interviewing, and other evidence about relevance, comprehensiveness, and comprehensibility.

## One workflow API

The three recommended workflows now share a stable object contract. A fitted
`sort_validity()`, `rating_validity()`, or `expert_validity()` object always
contains:

- `results` — item/result-level evidence;
- `scale_summary` — target-scale or panel-level evidence;
- `settings` — analysis choices;
- `design` — sample-size, missingness, and design metadata; and
- `details` — method-specific supporting results.

Every `results` table also includes a common `status` field with the restrained
categories **Supported**, **Review**, **Insufficient data**, or **Descriptive
only**. Method-specific `recommendation` wording is retained alongside it—for
example, item-sort and construct-rating workflows still use `Retain` when their
full statistical screening criterion is met. This keeps the methods faithful to
their evidentiary role while making programmatic use consistent across workflows.

`summary()` uses the same common count fields across all three workflows, and
`print()`/`plot()` retain method-appropriate displays. Compatibility aliases such
as `rating_fit$contrasts` and `expert_fit$scale` remain available for code written
before v0.0.6.

## Installation

```{r installation, eval=FALSE}
# install.packages("remotes")
remotes::install_github("JUhalt/contentvalidR")
```

## Recommended item-sort workflow

```{r sort-workflow}
library(contentvalidR)

sort_dat <- data.frame(
  item = rep(c("Clear 1", "Clear 2", "Needs review"), each = 20),
  rater = rep(1:20, 3),
  target_construct = "A",
  assigned_construct = c(
    rep("A", 18), rep("B", 2),
    rep("A", 16), rep("B", 4),
    rep("A", 12), rep("B", 8)
  )
)

fit <- sort_validity(sort_dat)
fit
summary(fit)
```

The workflow deliberately separates two levels of evidence:

- **Item level:** Psa/Csv plus the exact Howard-Melloy target-count rule produce **Retain**, **Review**, or **Insufficient data** flags. The output also names the strongest competing construct so a weak item is diagnostically useful rather than just "non-significant."
- **Target-scale level:** Psa and Csv are averaged across the target scale's items and interpreted using Colquitt et al. (2019)'s empirical percentile norms for definitional correspondence and distinctiveness. This matches how those norms were constructed.

`Review` deliberately does **not** mean automatic deletion. Likewise, Colquitt categories such as *Strong* or *Moderate* are empirical normative standing, not universal pass/fail cutoffs.

### Correlation-conditional Colquitt norms

If substantive data provide the average correlation between a focal scale and its orbiting scales, the workflow can select Colquitt et al.'s correlation-conditional norm set:

```{r colquitt-norms}
sort_validity(sort_dat, orbiting_r = .42)$scale_summary
colquitt_benchmarks("csv", orbiting_r = .42)
```

The published norms were developed with **naive judges representative of the target population**. If the sort used expert judges, declare that explicitly; the package will suppress Colquitt labels rather than apply an unsupported benchmark:

```{r expert-sort}
sort_validity(sort_dat, judge_type = "expert")$scale_summary
```

### Exact planning rather than a judge-count rule of thumb

```{r sort-planning}
sort_power(N = c(20, 30, 40), true_p = c(.60, .70, .80))
```

`sort_power()` gives the exact probability of reaching the Howard-Melloy retention count under each planned `N` and assumed true target-assignment probability.

### Low-level statistics

Researchers who need the component statistics directly can still use:

```{r sort-components}
compute_psa(sort_dat)
compute_csv(sort_dat)
csv_binom_test(n_c = 15, N = 20)
interpret_colquitt(.70, "csv")
```

Missing assignments are excluded itemwise and are reported explicitly in `n_missing` so the effective denominator is visible.


## Recommended construct-rating workflow

In the Hinkin-Tracey design, the **same judge rates each item against every construct definition**. `rating_validity()` treats that dependence explicitly rather than analyzing the ratings as independent groups.

```{r rating-workflow}
set.seed(12)
rating_dat <- expand.grid(
  item = c("A1", "A2", "B1"),
  rater = 1:20,
  construct = c("A", "B", "C")
)
rating_dat$target_construct <- ifelse(rating_dat$item == "B1", "B", "A")
rating_dat$rating <- ifelse(
  rating_dat$construct == rating_dat$target_construct,
  pmin(5, pmax(1, round(rnorm(nrow(rating_dat), 4.4, .6)))),
  pmin(5, pmax(1, round(rnorm(nrow(rating_dat), 2.1, .7))))
)

rfit <- rating_validity(rating_dat, scale_min = 1, scale_max = 5)
rfit
summary(rfit)
```

The workflow combines two descriptive indices with direct item-level screening:

- **HTC** (Hinkin-Tracey correspondence): how strongly ratings match the intended definition;
- **HTD** (Hinkin-Tracey distinctiveness): how much intended-definition ratings exceed orbiting-definition ratings;
- a **one-way repeated-measures ANOVA** for each item; and
- **planned paired contrasts** comparing the target definition with each orbiting definition.

As with the item-sort workflow, `Retain` and `Review` are screening labels rather than automatic editorial decisions. The output names the strongest orbiting competitor so a weak item tells the researcher *where* the conceptual overlap appears. Colquitt HTC/HTD labels are applied to target-scale averages, not treated as universal item-level cutoffs.

Low-level components remain available:

```{r rating-components}
htc(rating_dat, scale_min = 1, scale_max = 5)
htd(rating_dat, scale_min = 1, scale_max = 5)
anova_content(rating_dat)
```

## Recommended expert-panel workflow

Expert panels answer several *different* questions, so `expert_validity()` uses an explicit mode rather than pretending that Aiken V, CVR, CVI, and IOC are interchangeable.

### Relevance: Aiken V + CVI / modified kappa

```{r expert-relevance}
expert_ratings <- matrix(
  c(4,4,4,4,4,4,
    4,4,4,3,4,4,
    4,3,4,4,3,4),
  nrow = 6,
  dimnames = list(NULL, c("Item1", "Item2", "Item3"))
)

efit <- expert_validity(
  expert_ratings,
  mode = "relevance",
  lo = 1, hi = 4
)
efit
summary(efit)
```

Relevance mode reports Aiken's V with the Penfield-Giacobbi score confidence interval, I-CVI, Polit-Beck-Owen modified kappa, S-CVI/Ave, and S-CVI/UA. The workflow displays common panel-size CVI guidelines as **review aids**, not universal validity cutoffs. Aiken V is not converted into an automatic deletion rule.

The CVI relevance threshold is explicit and can be changed when a study uses a different rating convention:

```{r expert-relevance-cut}
expert_validity(expert_ratings, mode = "relevance",
                lo = 1, hi = 4, relevance_cut = 3)
```

### Essentiality: Lawshe CVR + exact inference

```{r expert-essentiality}
expert_validity(
  c(10, 8, 6),
  mode = "essentiality",
  N = 12
)
```

The CVR workflow derives the item-specific critical essential count directly from the exact binomial distribution, following the logic revisited by Ayre and Scally (2014). Judge-by-item 0/1 matrices are also accepted, including itemwise missingness when explicitly requested.

### Item-objective congruence

```{r expert-congruence}
ioc_dat <- expand.grid(
  item = c("I1", "I2"),
  judge = 1:4,
  objective = c("A", "B")
)
ioc_dat$target_objective <- ifelse(ioc_dat$item == "I1", "A", "B")
ioc_dat$score <- ifelse(
  ioc_dat$objective == ioc_dat$target_objective, 1, -1
)

expert_validity(ioc_dat, mode = "congruence")
```

When a target objective is supplied, the workflow reports the intended IOC, strongest competing objective, and target-minus-competitor margin. Without a target mapping, IOC cells are returned descriptively instead of manufacturing a pass/fail claim.

Low-level functions remain available for researchers who need the components directly:

```{r expert-components}
aikens_v(expert_ratings, lo = 1, hi = 4)
cvr(essential = c(8, 10, 5), N = 12)
cvi(expert_ratings >= 3)
ioc(ioc_dat[c("item", "judge", "objective", "score")])
```

## Interpretive visualization

The workflow objects include dependency-free base-R graphics designed around the
substantive questions in each method:

```{r visualization, fig.width=7, fig.height=5}
plot(fit, type = "map")
plot(rfit, type = "map")
plot(rfit, type = "profile")
plot(efit)
```

The sort and rating maps jointly display **definitional correspondence** and
**definitional distinctiveness**, with target-scale means distinguished from item
points. The rating profile plot shows the intended-definition mean against the
strongest competitor for every item. Expert-panel plots use Aiken score intervals,
panel-specific CVR criteria, or target-versus-competitor IOC gaps as appropriate.
The plots intentionally avoid converting scale-level empirical norms into item-level
cutoffs.

```{r visualization-power, fig.width=7, fig.height=4}
plot(sort_power(N = seq(10, 50, by = 5), true_p = c(.60, .70, .80)))
```

## Reproducible examples and reporting

The package ships five deterministic, human-readable CSV examples covering the
item-sort, construct-rating, relevance, essentiality, and IOC/congruence input
shapes. They are installed under `inst/extdata` and are regenerated from the
base-R provenance script in `data-raw/build-example-data.R`. This keeps the
worked examples inspectable outside R as well as reproducible inside the package.

For example:

```{r bundled-example, eval=FALSE}
sort_path <- system.file("extdata", "sort_example.csv", package = "contentvalidR")
bundled_sort <- utils::read.csv(sort_path, stringsAsFactors = FALSE)
sort_validity(bundled_sort)
```

See `vignette("reporting-examples", package = "contentvalidR")` for conservative
manuscript-ready methods/results scaffolds, table-building examples, and a
minimum reproducibility statement. Package citation metadata are available with
`citation("contentvalidR")`; the method bibliography is installed as
`REFERENCES.bib`.

## Auxiliary modules

The diagnostic, simulation, and Q-factor helpers remain available as auxiliary/compatibility functions, but they are not release-defining workflows for v0.1.0. The three recommended primary workflows are `sort_validity()`, `rating_validity()`, and `expert_validity()`. Their existing low-level function names are retained through the first public release; v0.0.6 introduces no gratuitous renaming or deprecation.

## Core methodological references

- Anderson, J. C., & Gerbing, D. W. (1991). Predicting the performance of measures in a confirmatory factor analysis with a pretest assessment of their substantive validities. *Journal of Applied Psychology, 76*(5), 732–740. https://doi.org/10.1037/0021-9010.76.5.732
- Howard, M. C., & Melloy, R. C. (2016). Evaluating item-sort task methods: The presentation of a new statistical significance formula and methodological best practices. *Journal of Business and Psychology, 31*(1), 173–186. https://doi.org/10.1007/s10869-015-9404-y
- Colquitt, J. A., Sabey, T. B., Rodell, J. B., & Hill, E. T. (2019). Content validation guidelines: Evaluation criteria for definitional correspondence and definitional distinctiveness. *Journal of Applied Psychology, 104*(10), 1243–1265. https://doi.org/10.1037/apl0000406
- Hinkin, T. R., & Tracey, J. B. (1999). An analysis of variance approach to content validation. *Organizational Research Methods, 2*(2), 175–186. https://doi.org/10.1177/109442819922004
- Polit, D. F., Beck, C. T., & Owen, S. V. (2007). Is the CVI an acceptable indicator of content validity? Appraisal and recommendations. *Research in Nursing & Health, 30*(4), 459–467. https://doi.org/10.1002/nur.20199
- Aiken, L. R. (1980). Content validity and reliability of single items or questionnaires. *Educational and Psychological Measurement, 40*(4), 955–959. https://doi.org/10.1177/001316448004000419
- Penfield, R. D., & Giacobbi, P. R., Jr. (2004). Applying a score confidence interval to Aiken's item content-relevance index. *Measurement in Physical Education and Exercise Science, 8*(4), 213–225. https://doi.org/10.1207/S15327841MPEE0804_3
- Lawshe, C. H. (1975). A quantitative approach to content validity. *Personnel Psychology, 28*(4), 563–575. https://doi.org/10.1111/j.1744-6570.1975.tb01393.x
- Ayre, C., & Scally, A. J. (2014). Critical values for Lawshe's content validity ratio: Revisiting the original methods of calculation. *Measurement and Evaluation in Counseling and Development, 47*(1), 79–86. https://doi.org/10.1177/0748175613513808
- Rovinelli, R. J., & Hambleton, R. K. (1977). On the use of content specialists in the assessment of criterion-referenced test item validity. *Dutch Journal of Educational Research, 2*, 49–60.
- Turner, R. C., & Carlson, L. (2003). Indexes of item-objective congruence for multidimensional items. *International Journal of Testing, 3*(2), 163–171. https://doi.org/10.1207/S15327574IJT0302_5
