---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
set.seed(1)
if (!requireNamespace("contentvalidR", quietly = TRUE)) {
  devtools::load_all(quiet = TRUE)
}
```

# contentvalidR

<!-- badges: start -->
[![R-CMD-check](https://github.com/JUhalt/contentvalidR/actions/workflows/R-CMD-check.yaml/badge.svg)](https://github.com/JUhalt/contentvalidR/actions/workflows/R-CMD-check.yaml)
[![License: GPL v3 only](https://img.shields.io/badge/License-GPL--3.0--only-blue.svg)](https://github.com/JUhalt/contentvalidR/blob/master/LICENSE.md)
[![Lifecycle: experimental](https://img.shields.io/badge/lifecycle-experimental-orange.svg)](https://lifecycle.r-lib.org/articles/stages.html)
<!-- badges: end -->

**Current stable release: 0.5.0.**

Development plans: [Roadmap](https://github.com/JUhalt/contentvalidR/blob/master/ROADMAP.md) · [Future research](https://github.com/JUhalt/contentvalidR/milestone/3) · [Open issues](https://github.com/JUhalt/contentvalidR/issues).
Documentation: [Package website](https://juhalt.github.io/contentvalidR/) · [Changelog](https://juhalt.github.io/contentvalidR/news/index.html).

**contentvalidR** provides quantitative tools for substantive and content-oriented scale pretesting. The package provides six complementary workflows:

1. **Item sorting** — Anderson & Gerbing (1991) Psa/Csv, exact item-level inference following Howard & Melloy (2016), and scale-level empirical norms from Colquitt et al. (2019).
2. **Construct ratings** — Hinkin & Tracey (1999) fully crossed ratings, HTC/HTD, Greenhouse-Geisser-aware repeated-measures item screening, and Colquitt et al. (2019) scale-level norms.
3. **Expert panels** — Aiken's V, Lawshe CVR, CVI/modified kappa with score intervals, item-objective congruence, and panel-level agreement.
4. **Delphi rounds** — consensus and stability across successive expert rounds: the share of experts agreeing against a threshold fixed before the study, and between-round weighted kappa following Holey et al. (2007), with the published alternatives selectable.
5. **Judge heterogeneity** — generalizability theory for content-validity ratings following Crocker, Llabre & Miller (1988), judge severity and rater effects, and leave-one-judge-out influence diagnostics.
6. **Domain coverage** — blueprint coverage, and expert-perceived content structure via the multidimensional scaling and clustering procedure of Sireci & Geisinger (1992).

The first four ask whether each **item** behaves as intended, or, in a Delphi study, whether the panel has settled on it. The last two ask questions no item-level index can reach: whether your conclusions depend on the particular **judges** you recruited, and whether your item set actually covers the **domain** you set out to measure. An item can only be rated if someone wrote it, so a perfect relevance index says nothing about the facet you forgot.

The design goal is **interpretable output rather than coefficient dumps**. Recommended workflow functions summarize what the evidence supports, flag items that need attention, and distinguish statistical screening from substantive decisions. Printed output defines every index it reports, so results can be read without first consulting the source papers; see [`vignette("reading-output")`](https://juhalt.github.io/contentvalidR/articles/reading-output.html).

> Quantitative content-validity statistics are one part of a broader validity argument. They complement, rather than replace, construct definition, domain coverage, qualitative expert feedback, cognitive interviewing, and other evidence about relevance, comprehensiveness, and comprehensibility.

## One workflow API

All six recommended workflows share a stable object contract. A fitted
`sort_validity()`, `rating_validity()`, `expert_validity()`, `delphi_validity()`,
`judge_validity()`, or `domain_validity()` object always contains:

- `results` — evidence at the workflow's unit of analysis;
- `scale_summary` — target-scale or panel-level evidence;
- `settings` — analysis choices;
- `design` — sample-size, missingness, and design metadata; and
- `details` — method-specific supporting results.

The unit of analysis in `results` differs by workflow, which matters when writing
code against them: `sort_validity()`, `rating_validity()`, `expert_validity()`,
and `delphi_validity()` return one row per **item**, `judge_validity()` one row
per **judge**, and `domain_validity()` one row per **blueprint cell**.

Every `results` table also includes a common `status` field with the restrained
categories **Supported**, **Review**, **Insufficient data**, or **Descriptive
only**. Method-specific `recommendation` wording is retained alongside it—for
example, item-sort and construct-rating workflows still use `Retain` when their
full statistical screening criterion is met. This keeps the methods faithful to
their evidentiary role while making programmatic use consistent across workflows.

`summary()` uses the same common count fields across every workflow, and
`print()`/`plot()` retain method-appropriate displays. Compatibility aliases such
as `rating_fit$contrasts` and `expert_fit$scale` remain available for code written
before v0.0.6.

## Installation

Install the current stable release from the JUhalt R-universe:

```{r installation-stable, eval=FALSE}
install.packages(
  "contentvalidR",
  repos = c(
    "https://juhalt.r-universe.dev",
    "https://cloud.r-project.org"
  )
)
```

Install the current development version directly from GitHub:

```{r installation-development, eval=FALSE}
# install.packages("remotes")
remotes::install_github("JUhalt/contentvalidR")
```
## Recommended item-sort workflow

```{r sort-workflow}
library(contentvalidR)

sort_dat <- data.frame(
  item = rep(c("Clear 1", "Clear 2", "Needs review"), each = 20),
  rater = rep(1:20, 3),
  target_construct = "A",
  assigned_construct = c(
    rep("A", 18), rep("B", 2),
    rep("A", 16), rep("B", 4),
    rep("A", 12), rep("B", 8)
  )
)

fit <- sort_validity(sort_dat)
fit
summary(fit)
```

The workflow deliberately separates two levels of evidence:

- **Item level:** Psa/Csv plus the exact Howard-Melloy target-count rule produce **Retain**, **Review**, or **Insufficient data** flags. Psa is reported with an interval (`psa_low`, `psa_high`), so an item sorted by few judges does not look more settled than it is. The output also names the strongest competing construct so a weak item is diagnostically useful rather than just "non-significant."
- **Target-scale level:** Psa and Csv are averaged across the target scale's items and interpreted using Colquitt et al. (2019)'s empirical percentile norms for definitional correspondence and distinctiveness. This matches how those norms were constructed.

`Review` deliberately does **not** mean automatic deletion. Likewise, Colquitt categories such as *Strong* or *Moderate* are empirical normative standing, not universal pass/fail cutoffs.

### Correlation-conditional Colquitt norms

If substantive data provide the average correlation between a focal scale and its orbiting scales, the workflow can select Colquitt et al.'s correlation-conditional norm set:

```{r colquitt-norms}
sort_validity(sort_dat, orbiting_r = .42)$scale_summary
colquitt_benchmarks("csv", orbiting_r = .42)
```

The published norms were developed with **naive judges representative of the target population**. If the sort used expert judges, declare that explicitly; the package will suppress Colquitt labels rather than apply an unsupported benchmark:

```{r expert-sort}
sort_validity(sort_dat, judge_type = "expert")$scale_summary
```

### Exact planning rather than a judge-count rule of thumb

```{r sort-planning}
sort_power(N = c(20, 30, 40), true_p = c(.60, .70, .80))
```

`sort_power()` gives the exact probability of reaching the Howard-Melloy retention count under each planned `N` and assumed true target-assignment probability.

### Low-level statistics

Researchers who need the component statistics directly can still use:

```{r sort-components}
compute_psa(sort_dat)
compute_csv(sort_dat)
csv_binom_test(n_c = 15, N = 20)
interpret_colquitt(.70, "csv")
```

Missing assignments are excluded itemwise and are reported explicitly in `n_missing` so the effective denominator is visible.


## Recommended construct-rating workflow

In the Hinkin-Tracey design, the **same judge rates each item against every construct definition**. `rating_validity()` treats that dependence explicitly rather than analyzing the ratings as independent groups.

```{r rating-workflow}
set.seed(12)
rating_dat <- expand.grid(
  item = c("A1", "A2", "B1"),
  rater = 1:20,
  construct = c("A", "B", "C")
)
rating_dat$target_construct <- ifelse(rating_dat$item == "B1", "B", "A")
rating_dat$rating <- ifelse(
  rating_dat$construct == rating_dat$target_construct,
  pmin(5, pmax(1, round(rnorm(nrow(rating_dat), 4.4, .6)))),
  pmin(5, pmax(1, round(rnorm(nrow(rating_dat), 2.1, .7))))
)

rfit <- rating_validity(rating_dat, scale_min = 1, scale_max = 5)
rfit
summary(rfit)
```

The workflow combines two descriptive indices with direct item-level screening:

- **HTC** (Hinkin-Tracey correspondence): how strongly ratings match the intended definition;
- **HTD** (Hinkin-Tracey distinctiveness): how much intended-definition ratings exceed orbiting-definition ratings;
- a **one-way repeated-measures ANOVA** for each item; and
- **planned paired contrasts** comparing the target definition with each orbiting definition.

As with the item-sort workflow, `Retain` and `Review` are screening labels rather than automatic editorial decisions. The output names the strongest orbiting competitor so a weak item tells the researcher *where* the conceptual overlap appears. Colquitt HTC/HTD labels are applied to target-scale averages, not treated as universal item-level cutoffs.

Low-level components remain available:

```{r rating-components}
htc(rating_dat, scale_min = 1, scale_max = 5)
htd(rating_dat, scale_min = 1, scale_max = 5)
anova_content(rating_dat)
```

## Recommended expert-panel workflow

Expert panels answer several *different* questions, so `expert_validity()` uses an explicit mode rather than pretending that Aiken V, CVR, CVI, and IOC are interchangeable.

### Relevance: Aiken V + CVI / modified kappa

```{r expert-relevance}
expert_ratings <- matrix(
  c(4,4,4,4,4,4,
    4,4,4,3,4,4,
    4,3,4,4,3,4),
  nrow = 6,
  dimnames = list(NULL, c("Item1", "Item2", "Item3"))
)

efit <- expert_validity(
  expert_ratings,
  mode = "relevance",
  lo = 1, hi = 4,
  seed = 1
)
efit
summary(efit)
```

Relevance mode reports Aiken's V with the Penfield-Giacobbi score confidence interval, I-CVI with its own interval, Polit-Beck-Owen modified kappa, S-CVI/Ave, S-CVI/UA, and panel-level agreement. The workflow displays common panel-size CVI guidelines as **review aids**, not universal validity cutoffs. Aiken V is not converted into an automatic deletion rule.

Panel-level agreement is one coefficient for how consistently the panel rated the whole item set; modified kappa still describes each item. The default is Krippendorff's alpha on the ordinal ratings (Hayes & Krippendorff, 2007), and `agreement_level` switches to nominal or interval. Its bootstrap interval resamples items with all of their ratings, following Zapf et al. (2016); set `seed` to make it reproducible. In the example above, alpha is near zero even though most rating pairs are identical: the three items were rated almost the same, so the ratings barely distinguish one item from another. That is why the output reports identical rating pairs beside alpha. Gwet's AC1 is available through `agreement = "ac1"` but is never the default, and its output repeats the critique in Vach and Gerke (2023). `panel_agreement()` runs the same analysis on any rater-by-item matrix.

I-CVI here, like Psa in `sort_validity()`, is a proportion of a small panel, so it comes with an interval. The Wilson score interval is the default, following Newcombe (1998). Agresti-Coull and Clopper-Pearson exact intervals are available through `proportion_ci` for studies that need to match earlier work, and the printed output names whichever method ran:

```{r expert-relevance-interval}
exact_fit <- expert_validity(expert_ratings, mode = "relevance",
                             lo = 1, hi = 4, proportion_ci = "exact")
exact_fit$results[, c("item", "I_CVI", "I_CVI_low", "I_CVI_high")]
```

The CVI relevance threshold is explicit and can be changed when a study uses a different rating convention:

```{r expert-relevance-cut}
expert_validity(expert_ratings, mode = "relevance",
                lo = 1, hi = 4, relevance_cut = 3, seed = 1)
```

### Essentiality: Lawshe CVR + exact inference

```{r expert-essentiality}
expert_validity(
  c(10, 8, 6),
  mode = "essentiality",
  N = 12
)
```

The CVR workflow derives the item-specific critical essential count directly from the exact binomial distribution, following the logic revisited by Ayre and Scally (2014). Judge-by-item 0/1 matrices are also accepted, including itemwise missingness when explicitly requested.

### Item-objective congruence

```{r expert-congruence}
ioc_dat <- expand.grid(
  item = c("I1", "I2"),
  judge = 1:4,
  objective = c("A", "B")
)
ioc_dat$target_objective <- ifelse(ioc_dat$item == "I1", "A", "B")
ioc_dat$score <- ifelse(
  ioc_dat$objective == ioc_dat$target_objective, 1, -1
)

expert_validity(ioc_dat, mode = "congruence")
```

When a target objective is supplied, the workflow reports the intended IOC, strongest competing objective, and target-minus-competitor margin. Without a target mapping, IOC cells are returned descriptively instead of manufacturing a pass/fail claim.

Low-level functions remain available for researchers who need the components directly:

```{r expert-components}
aikens_v(expert_ratings, lo = 1, hi = 4)
cvr(essential = c(8, 10, 5), N = 12)
cvi(expert_ratings >= 3)
ioc(ioc_dat[c("item", "judge", "objective", "score")])
```

## From content validity to empirical validation

Content-validity evidence ends where response data begin. `content_handoff()` packages a finished workflow's item decisions — the items that survived review, a per-item evidence table, and the provenance of the analysis — so the item set and its reasons travel together instead of being retyped:

```{r handoff}
handoff <- content_handoff(efit)
handoff$items
handoff$item_evidence[, c("item", "carried", "status", "n_judges", "rule")]
```

The object carries the carried item names, a construct mapping (`scales`, when the design has one), a per-item evidence table, the statistics behind each decision, and the provenance of the analysis. Only items with a `Supported` status travel by default; `keep` widens that when a protocol carries items flagged for review. Items held back stay in the table with `carried = FALSE`, because review is not deletion.

Once responses are collected, the item set carries into the empirical stage. In [nomologR](https://github.com/JUhalt/nomologR), the companion package for that stage, that is `nomo_screen(responses, items = handoff)`. The object shape is agreed between the two packages as schema version 1, and every field is a base type, so neither package depends on the other. Carrying an item forward is not a prediction that it will perform: a clearly relevant item can still correlate poorly with its construct or load on an unintended factor, which is what the empirical analysis tests. See `vignette("handoff-to-empirical-validation")`.

## Recommended judge-heterogeneity workflow

Aggregate indices average heterogeneity away. `judge_validity()` asks whether your conclusions depend on the particular judges who happened to serve, and returns **one row per judge**.

```{r judge-workflow}
judge_ratings <- rbind(
  c(4, 4, 4, 3, 2, 2), c(4, 4, 3, 4, 2, 1), c(4, 3, 4, 4, 1, 2),
  c(3, 4, 4, 4, 2, 2), c(4, 4, 4, 4, 2, 1), c(4, 3, 4, 3, 1, 2),
  c(4, 4, 3, 4, 2, 2), c(2, 2, 2, 2, 1, 1)
)
dimnames(judge_ratings) <- list(paste0("Judge", 1:8), paste0("Item", 1:6))

judge_fit <- judge_validity(judge_ratings, lo = 1, hi = 4)
judge_fit$results[, c("judge", "mean_rating", "severity_raw",
                      "differentiation", "n_items_flipped", "recommendation")]
```

Severity is signed so positive means harsher. `n_items_flipped` is the influence diagnostic: how many items would change review status if that judge were removed. A judge flagged here is **not** a judge to delete — a dissenting expert may be the one reading the construct definition correctly.

Generalizability theory answers the planning question of how many judges the design actually needs:

```{r judge-gtheory}
gt <- gtheory_content(judge_ratings)
gt$coefficients
gt$judges_needed
```

The dependability coefficient (`phi_coefficient`) concerns the absolute level of ratings and is penalized by judge severity differences, which is usually what content-validity decisions rest on. `NA` in `judges_needed` means no realistic panel reaches that target, which happens when judges barely distinguished the items.

## Recommended domain-coverage workflow

Relevance indices describe items that exist. They cannot reveal a facet nobody wrote an item for. `domain_validity()` returns **one row per blueprint cell**.

```{r domain-workflow}
assignments <- data.frame(
  item = paste0("I", 1:7),
  construct = c("Autonomy", "Autonomy", "Autonomy", "Autonomy",
                "Competence", "Competence", "Relatedness"),
  stringsAsFactors = FALSE
)

domain_fit <- domain_validity(
  assignments,
  cell_col = "construct",
  domain = c("Autonomy", "Competence", "Relatedness", "Belonging")
)
domain_fit$results[, c("cell", "n_items", "share", "recommendation")]
```

`Belonging` is the point: the blueprint asks for it and nothing addresses it. Detecting that requires passing the full cell list through `domain`, since an empty cell leaves no trace in the item assignments. Omit it and the output says coverage gaps could not be detected rather than implying complete coverage.

Where experts rated item similarity, `content_structure()` tests whether they group items the way the blueprint claims, using multidimensional scaling and clustering with a chance-corrected adjusted Rand index. `similarity_from_sort()` derives those similarities from an existing sorting task.

## Planning, rounds, and reporting

`expert_power()` replaces "use about six experts" with a question that has an answer:

```{r expert-planning}
expert_power(n_experts = 3:8, prob = 0.9)$results
```

Note the step. The common I-CVI guideline requires unanimity up to five experts and 0.78 from six, so a fourth or fifth expert **lowers** the probability of clearing while a sixth raises it sharply. That is a property of the guideline, not of the items, and the package reports it rather than smoothing it away.

`compare_rounds()` compares successive pretest rounds and, critically, checks whether the analysis settings changed between them — so a relaxed criterion cannot read as item improvement.

For a Delphi study, where the same panel rates items over several rounds with feedback in between, `delphi_validity()` reports two questions separately. Consensus is the share of experts agreeing, judged against a threshold you fix before the study. Stability is the share of experts who kept their rating, shown beside weighted kappa between rounds. Kappa carries no verbal labels, because it falls as a panel converges. The published alternatives are selectable, each with its limits printed: Chaffin and Talley's lambda and chi-square, Dajani et al.'s chi-square, and Scheibe et al.'s 15% rule. See `vignette("delphi-rounds")`.

`content_report()` builds manuscript-ready tables as a data frame or as Markdown for Quarto and R Markdown, with no reporting dependency added to the package.

```{r reporting}
content_report(fit, include = "flagged")
```

There is deliberately no helper returning "the items that passed." Filtering on `status` is a substantive decision that belongs in your own visible code.

## Interpretive visualization

The workflow objects include dependency-free base-R graphics designed around the
substantive questions in each method:

```{r visualization, fig.width=7, fig.height=5}
plot(fit, type = "map")
plot(rfit, type = "map")
plot(rfit, type = "profile")
plot(efit)
```

The sort and rating maps jointly display **definitional correspondence** and
**definitional distinctiveness**, with target-scale means distinguished from item
points. The rating profile plot shows the intended-definition mean against the
strongest competitor for every item. Expert-panel plots use Aiken score intervals,
panel-specific CVR criteria, or target-versus-competitor IOC gaps as appropriate.
The plots intentionally avoid converting scale-level empirical norms into item-level
cutoffs.

```{r visualization-power, fig.width=7, fig.height=4}
plot(sort_power(N = seq(10, 50, by = 5), true_p = c(.60, .70, .80)))
```

## Reproducible examples and reporting

The package ships five deterministic, human-readable CSV examples covering the
item-sort, construct-rating, relevance, essentiality, and IOC/congruence input
shapes. They are installed under `inst/extdata` and are regenerated from the
base-R provenance script in `data-raw/build-example-data.R`. This keeps the
worked examples inspectable outside R as well as reproducible inside the package.

For example:

```{r bundled-example, eval=FALSE}
sort_path <- system.file("extdata", "sort_example.csv", package = "contentvalidR")
bundled_sort <- utils::read.csv(sort_path, stringsAsFactors = FALSE)
sort_validity(bundled_sort)
```

See `vignette("reporting-examples", package = "contentvalidR")` for conservative
manuscript-ready methods/results scaffolds, table-building examples, and a
minimum reproducibility statement. Package citation metadata are available with
`citation("contentvalidR")`; the method bibliography is installed as
`REFERENCES.bib`.

## Auxiliary modules

The diagnostic, simulation, and Q-factor helpers remain available as auxiliary functions, but they are not recommended workflows. The six recommended workflows are `sort_validity()`, `rating_validity()`, `expert_validity()`, `delphi_validity()`, `judge_validity()`, and `domain_validity()`.

## Core methodological references

- Anderson, J. C., & Gerbing, D. W. (1991). Predicting the performance of measures in a confirmatory factor analysis with a pretest assessment of their substantive validities. *Journal of Applied Psychology, 76*(5), 732–740. https://doi.org/10.1037/0021-9010.76.5.732
- Howard, M. C., & Melloy, R. C. (2016). Evaluating item-sort task methods: The presentation of a new statistical significance formula and methodological best practices. *Journal of Business and Psychology, 31*(1), 173–186. https://doi.org/10.1007/s10869-015-9404-y
- Colquitt, J. A., Sabey, T. B., Rodell, J. B., & Hill, E. T. (2019). Content validation guidelines: Evaluation criteria for definitional correspondence and definitional distinctiveness. *Journal of Applied Psychology, 104*(10), 1243–1265. https://doi.org/10.1037/apl0000406
- Hinkin, T. R., & Tracey, J. B. (1999). An analysis of variance approach to content validation. *Organizational Research Methods, 2*(2), 175–186. https://doi.org/10.1177/109442819922004
- Polit, D. F., Beck, C. T., & Owen, S. V. (2007). Is the CVI an acceptable indicator of content validity? Appraisal and recommendations. *Research in Nursing & Health, 30*(4), 459–467. https://doi.org/10.1002/nur.20199
- Aiken, L. R. (1980). Content validity and reliability of single items or questionnaires. *Educational and Psychological Measurement, 40*(4), 955–959. https://doi.org/10.1177/001316448004000419
- Penfield, R. D., & Giacobbi, P. R., Jr. (2004). Applying a score confidence interval to Aiken's item content-relevance index. *Measurement in Physical Education and Exercise Science, 8*(4), 213–225. https://doi.org/10.1207/S15327841MPEE0804_3
- Lawshe, C. H. (1975). A quantitative approach to content validity. *Personnel Psychology, 28*(4), 563–575. https://doi.org/10.1111/j.1744-6570.1975.tb01393.x
- Ayre, C., & Scally, A. J. (2014). Critical values for Lawshe's content validity ratio: Revisiting the original methods of calculation. *Measurement and Evaluation in Counseling and Development, 47*(1), 79–86. https://doi.org/10.1177/0748175613513808
- Rovinelli, R. J., & Hambleton, R. K. (1977). On the use of content specialists in the assessment of criterion-referenced test item validity. *Dutch Journal of Educational Research, 2*, 49–60.
- Turner, R. C., & Carlson, L. (2003). Indexes of item-objective congruence for multidimensional items. *International Journal of Testing, 3*(2), 163–171. https://doi.org/10.1207/S15327574IJT0302_5
- Crocker, L., Llabre, M., & Miller, M. D. (1988). The generalizability of content validity ratings. *Journal of Educational Measurement, 25*(4), 287–299. https://doi.org/10.1111/j.1745-3984.1988.tb00309.x
- Engelhard, G. (1994). Examining rater errors in the assessment of written composition with a many-faceted Rasch model. *Journal of Educational Measurement, 31*(2), 93–112. https://doi.org/10.1111/j.1745-3984.1994.tb00436.x
- Linacre, J. M. (1989). *Many-Facet Rasch Measurement.* MESA Press.
- de Boeck, P., & Wilson, M. (2004). *Explanatory Item Response Models: A Generalized Linear and Nonlinear Approach.* Springer.
- Sireci, S. G., & Geisinger, K. F. (1992). Analyzing test content using cluster analysis and multidimensional scaling. *Applied Psychological Measurement, 16*(1), 17–31. https://doi.org/10.1177/014662169201600102
- Sireci, S. G., & Geisinger, K. F. (1995). Using subject-matter experts to assess content representation: An MDS analysis. *Applied Psychological Measurement, 19*(3), 241–255. https://doi.org/10.1177/014662169501900303
- Sireci, S. G. (1998). The construct of content validity. *Social Indicators Research, 45*(1–3), 83–117. https://doi.org/10.1023/A:1006985528729
- Hubert, L., & Arabie, P. (1985). Comparing partitions. *Journal of Classification, 2*(1), 193–218. https://doi.org/10.1007/BF01908075
- Lynn, M. R. (1986). Determination and quantification of content validity. *Nursing Research, 35*(6), 382–385.
- Wilson, E. B. (1927). Probable inference, the law of succession, and statistical inference. *Journal of the American Statistical Association, 22*(158), 209–212. https://doi.org/10.1080/01621459.1927.10502953
- Clopper, C. J., & Pearson, E. S. (1934). The use of confidence or fiducial limits illustrated in the case of the binomial. *Biometrika, 26*(4), 404–413. https://doi.org/10.1093/biomet/26.4.404
- Agresti, A., & Coull, B. A. (1998). Approximate is better than "exact" for interval estimation of binomial proportions. *The American Statistician, 52*(2), 119–126. https://doi.org/10.1080/00031305.1998.10480550
- Newcombe, R. G. (1998). Two-sided confidence intervals for the single proportion: Comparison of seven methods. *Statistics in Medicine, 17*(8), 857–872.
- Hayes, A. F., & Krippendorff, K. (2007). Answering the call for a standard reliability measure for coding data. *Communication Methods and Measures, 1*(1), 77–89. https://doi.org/10.1080/19312450709336664
- Zapf, A., Castell, S., Morawietz, L., & Karch, A. (2016). Measuring inter-rater reliability for nominal data: Which coefficients and confidence intervals are appropriate? *BMC Medical Research Methodology, 16*, 93. https://doi.org/10.1186/s12874-016-0200-9
- Gwet, K. L. (2008). Computing inter-rater reliability and its variance in the presence of high agreement. *British Journal of Mathematical and Statistical Psychology, 61*(1), 29–48. https://doi.org/10.1348/000711006X126600
- Vach, W., & Gerke, O. (2023). Gwet's AC1 is not a substitute for Cohen's kappa: A comparison of basic properties. *MethodsX, 10*, 102212.

## License

contentvalidR is licensed under the **GNU General Public License, version 3
only** (SPDX: `GPL-3.0-only`; R metadata: `GPL-3`). See the
[full license](https://github.com/JUhalt/contentvalidR/blob/master/LICENSE.md). Copyright (c) 2025-2026 Joshua Uhalt.

Releases from v0.2.0 onward carry the GPLv3-only license. The earlier
[v0.1.0 release](https://github.com/JUhalt/contentvalidR/releases/tag/v0.1.0)
remains available under its original MIT license, and historical attribution is
preserved in
[inst/NOTICE](https://github.com/JUhalt/contentvalidR/blob/master/inst/NOTICE).
