---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
set.seed(1)
if (!requireNamespace("contentvalidR", quietly = TRUE)) {
  devtools::load_all(quiet = TRUE)
}
```

# contentvalidR

<!-- badges: start -->
[![R-CMD-check](https://github.com/JUhalt/contentvalidR/actions/workflows/R-CMD-check.yaml/badge.svg)](https://github.com/JUhalt/contentvalidR/actions/workflows/R-CMD-check.yaml)
[![License: MIT](https://img.shields.io/badge/License-MIT-yellow.svg)](LICENSE)
[![Lifecycle: experimental](https://img.shields.io/badge/lifecycle-experimental-orange.svg)](https://lifecycle.r-lib.org/articles/stages.html)
<!-- badges: end -->

**contentvalidR** provides quantitative tools for substantive and content-oriented scale pretesting. The package is being developed around three complementary workflows:

1. **Item sorting** — Anderson & Gerbing (1991) Psa/Csv with exact item-sort inference following Howard & Melloy (2016).
2. **Construct ratings** — Hinkin & Tracey (1999) rating-based procedures. This module remains experimental while the repeated-rating design is being modernized.
3. **Expert panels** — Aiken's V, Lawshe CVR, CVI/modified kappa, and item-objective congruence.

The design goal is **interpretable output rather than coefficient dumps**. Recommended workflow functions summarize what the evidence supports, flag items that need attention, and distinguish statistical screening from substantive decisions.

> Quantitative content-validity statistics are one part of a broader validity argument. They complement, rather than replace, construct definition, domain coverage, qualitative expert feedback, cognitive interviewing, and other evidence about relevance, comprehensiveness, and comprehensibility.

## Installation

```{r installation, eval=FALSE}
# install.packages("remotes")
remotes::install_github("JUhalt/contentvalidR")
```

## Recommended item-sort workflow

```{r sort-workflow}
library(contentvalidR)

sort_dat <- data.frame(
  item = rep(c("Clear item", "Needs review"), each = 20),
  rater = rep(1:20, 2),
  target_construct = rep("A", 40),
  assigned_construct = c(
    rep("A", 16), rep("B", 4),
    rep("A", 11), rep("B", 9)
  )
)

fit <- sort_validity(sort_dat)
fit
summary(fit)
```

`sort_validity()` combines:

- **Psa**: proportion assigned to the intended construct;
- **Csv**: intended assignments minus the most common competing assignment, divided by the effective judge count;
- an **exact target-count test** following Howard & Melloy (2016); and
- a restrained recommendation: **Retain**, **Review**, or **Insufficient data**.

`Review` deliberately does **not** mean automatic deletion. The researcher should examine wording, construct overlap, theoretical coverage, and qualitative feedback before revising or removing an item.

### Low-level statistics

Researchers who need the component statistics directly can still use:

```{r sort-components}
compute_psa(sort_dat)
compute_csv(sort_dat)
csv_binom_test(n_c = 15, N = 20)
```

Missing assignments are excluded itemwise and are reported explicitly in `n_missing` so the effective denominator is visible.

## Expert-panel indices

```{r expert-indices}
# Aiken's V: judges x items
R <- matrix(sample(1:5, 5 * 4, replace = TRUE), nrow = 5)
aikens_v(R, lo = 1, hi = 5)

# Lawshe CVR
cvr(essential = c(8, 10, 5), N = 12)

# CVI with Polit-Beck-Owen modified kappa
M <- matrix(
  c(1,1,1,1,
    1,1,1,0,
    1,1,0,0),
  nrow = 4,
  dimnames = list(NULL, c("Item1", "Item2", "Item3"))
)
cvi(M)
```

`cvi()` reports I-CVI, S-CVI/Ave, S-CVI/UA, the exact chance-agreement probability, and modified kappa. Its print method also reminds users that those indices do not, by themselves, establish the full content-validity argument.

## Experimental modules

The rating/ANOVA, diagnostic, simulation, and Q-factor helpers remain available while their APIs and methodological scope are being hardened. They should currently be treated as experimental rather than as the package's primary public interface.

## Core methodological references

- Anderson, J. C., & Gerbing, D. W. (1991). Predicting the performance of measures in a confirmatory factor analysis with a pretest assessment of their substantive validities. *Journal of Applied Psychology, 76*(5), 732–740. https://doi.org/10.1037/0021-9010.76.5.732
- Howard, M. C., & Melloy, R. C. (2016). Evaluating item-sort task methods: The presentation of a new statistical significance formula and methodological best practices. *Journal of Business and Psychology, 31*(1), 173–186. https://doi.org/10.1007/s10869-015-9404-y
- Hinkin, T. R., & Tracey, J. B. (1999). An analysis of variance approach to content validation. *Organizational Research Methods, 2*(2), 175–186.
- Polit, D. F., Beck, C. T., & Owen, S. V. (2007). Is the CVI an acceptable indicator of content validity? Appraisal and recommendations. *Research in Nursing & Health, 30*(4), 459–467. https://doi.org/10.1002/nur.20199
