---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
set.seed(1)
if (!requireNamespace("contentvalidR", quietly = TRUE)) {
  devtools::load_all(quiet = TRUE)
}
```

# contentvalidR

<!-- badges: start -->
[![R-CMD-check](https://github.com/JUhalt/contentvalidR/actions/workflows/R-CMD-check.yaml/badge.svg)](https://github.com/JUhalt/contentvalidR/actions/workflows/R-CMD-check.yaml)
[![License: MIT](https://img.shields.io/badge/License-MIT-yellow.svg)](LICENSE)
[![Lifecycle: experimental](https://img.shields.io/badge/lifecycle-experimental-orange.svg)](https://lifecycle.r-lib.org/articles/stages.html)
<!-- badges: end -->

**contentvalidR** provides quantitative tools for substantive and content-oriented scale pretesting. The package is being developed around three complementary workflows:

1. **Item sorting** — Anderson & Gerbing (1991) Psa/Csv, exact item-level inference following Howard & Melloy (2016), and scale-level empirical norms from Colquitt et al. (2019).
2. **Construct ratings** — Hinkin & Tracey (1999) fully crossed ratings, HTC/HTD, Greenhouse-Geisser-aware repeated-measures item screening, and Colquitt et al. (2019) scale-level norms.
3. **Expert panels** — Aiken's V, Lawshe CVR, CVI/modified kappa, and item-objective congruence.

The design goal is **interpretable output rather than coefficient dumps**. Recommended workflow functions summarize what the evidence supports, flag items that need attention, and distinguish statistical screening from substantive decisions.

> Quantitative content-validity statistics are one part of a broader validity argument. They complement, rather than replace, construct definition, domain coverage, qualitative expert feedback, cognitive interviewing, and other evidence about relevance, comprehensiveness, and comprehensibility.

## Installation

```{r installation, eval=FALSE}
# install.packages("remotes")
remotes::install_github("JUhalt/contentvalidR")
```

## Recommended item-sort workflow

```{r sort-workflow}
library(contentvalidR)

sort_dat <- data.frame(
  item = rep(c("Clear 1", "Clear 2", "Needs review"), each = 20),
  rater = rep(1:20, 3),
  target_construct = "A",
  assigned_construct = c(
    rep("A", 18), rep("B", 2),
    rep("A", 16), rep("B", 4),
    rep("A", 12), rep("B", 8)
  )
)

fit <- sort_validity(sort_dat)
fit
summary(fit)
```

The workflow deliberately separates two levels of evidence:

- **Item level:** Psa/Csv plus the exact Howard-Melloy target-count rule produce **Retain**, **Review**, or **Insufficient data** flags. The output also names the strongest competing construct so a weak item is diagnostically useful rather than just "non-significant."
- **Target-scale level:** Psa and Csv are averaged across the target scale's items and interpreted using Colquitt et al. (2019)'s empirical percentile norms for definitional correspondence and distinctiveness. This matches how those norms were constructed.

`Review` deliberately does **not** mean automatic deletion. Likewise, Colquitt categories such as *Strong* or *Moderate* are empirical normative standing, not universal pass/fail cutoffs.

### Correlation-conditional Colquitt norms

If substantive data provide the average correlation between a focal scale and its orbiting scales, the workflow can select Colquitt et al.'s correlation-conditional norm set:

```{r colquitt-norms}
sort_validity(sort_dat, orbiting_r = .42)$scale_summary
colquitt_benchmarks("csv", orbiting_r = .42)
```

The published norms were developed with **naive judges representative of the target population**. If the sort used expert judges, declare that explicitly; the package will suppress Colquitt labels rather than apply an unsupported benchmark:

```{r expert-sort}
sort_validity(sort_dat, judge_type = "expert")$scale_summary
```

### Exact planning rather than a judge-count rule of thumb

```{r sort-planning}
sort_power(N = c(20, 30, 40), true_p = c(.60, .70, .80))
```

`sort_power()` gives the exact probability of reaching the Howard-Melloy retention count under each planned `N` and assumed true target-assignment probability.

### Low-level statistics

Researchers who need the component statistics directly can still use:

```{r sort-components}
compute_psa(sort_dat)
compute_csv(sort_dat)
csv_binom_test(n_c = 15, N = 20)
interpret_colquitt(.70, "csv")
```

Missing assignments are excluded itemwise and are reported explicitly in `n_missing` so the effective denominator is visible.


## Recommended construct-rating workflow

In the Hinkin-Tracey design, the **same judge rates each item against every construct definition**. `rating_validity()` treats that dependence explicitly rather than analyzing the ratings as independent groups.

```{r rating-workflow}
set.seed(12)
rating_dat <- expand.grid(
  item = c("A1", "A2", "B1"),
  rater = 1:20,
  construct = c("A", "B", "C")
)
rating_dat$target_construct <- ifelse(rating_dat$item == "B1", "B", "A")
rating_dat$rating <- ifelse(
  rating_dat$construct == rating_dat$target_construct,
  pmin(5, pmax(1, round(rnorm(nrow(rating_dat), 4.4, .6)))),
  pmin(5, pmax(1, round(rnorm(nrow(rating_dat), 2.1, .7))))
)

rfit <- rating_validity(rating_dat, scale_min = 1, scale_max = 5)
rfit
summary(rfit)
```

The workflow combines two descriptive indices with direct item-level screening:

- **HTC** (Hinkin-Tracey correspondence): how strongly ratings match the intended definition;
- **HTD** (Hinkin-Tracey distinctiveness): how much intended-definition ratings exceed orbiting-definition ratings;
- a **one-way repeated-measures ANOVA** for each item; and
- **planned paired contrasts** comparing the target definition with each orbiting definition.

As with the item-sort workflow, `Retain` and `Review` are screening labels rather than automatic editorial decisions. The output names the strongest orbiting competitor so a weak item tells the researcher *where* the conceptual overlap appears. Colquitt HTC/HTD labels are applied to target-scale averages, not treated as universal item-level cutoffs.

Low-level components remain available:

```{r rating-components}
htc(rating_dat, scale_min = 1, scale_max = 5)
htd(rating_dat, scale_min = 1, scale_max = 5)
anova_content(rating_dat)
```

## Expert-panel indices

```{r expert-indices}
# Aiken's V: judges x items
R <- matrix(sample(1:5, 5 * 4, replace = TRUE), nrow = 5)
aikens_v(R, lo = 1, hi = 5)

# Lawshe CVR
cvr(essential = c(8, 10, 5), N = 12)

# CVI with Polit-Beck-Owen modified kappa
M <- matrix(
  c(1,1,1,1,
    1,1,1,0,
    1,1,0,0),
  nrow = 4,
  dimnames = list(NULL, c("Item1", "Item2", "Item3"))
)
cvi(M)
```

`cvi()` reports I-CVI, S-CVI/Ave, S-CVI/UA, the exact chance-agreement probability, and modified kappa. Its print method also reminds users that those indices do not, by themselves, establish the full content-validity argument.

## Experimental modules

The diagnostic, simulation, and Q-factor helpers remain available while their APIs and methodological scope are being hardened. The two recommended primary workflows are now `sort_validity()` and `rating_validity()`.

## Core methodological references

- Anderson, J. C., & Gerbing, D. W. (1991). Predicting the performance of measures in a confirmatory factor analysis with a pretest assessment of their substantive validities. *Journal of Applied Psychology, 76*(5), 732–740. https://doi.org/10.1037/0021-9010.76.5.732
- Howard, M. C., & Melloy, R. C. (2016). Evaluating item-sort task methods: The presentation of a new statistical significance formula and methodological best practices. *Journal of Business and Psychology, 31*(1), 173–186. https://doi.org/10.1007/s10869-015-9404-y
- Colquitt, J. A., Sabey, T. B., Rodell, J. B., & Hill, E. T. (2019). Content validation guidelines: Evaluation criteria for definitional correspondence and definitional distinctiveness. *Journal of Applied Psychology, 104*(10), 1243–1265. https://doi.org/10.1037/apl0000406
- Hinkin, T. R., & Tracey, J. B. (1999). An analysis of variance approach to content validation. *Organizational Research Methods, 2*(2), 175–186. https://doi.org/10.1177/109442819922004
- Polit, D. F., Beck, C. T., & Owen, S. V. (2007). Is the CVI an acceptable indicator of content validity? Appraisal and recommendations. *Research in Nursing & Health, 30*(4), 459–467. https://doi.org/10.1002/nur.20199
