---
title: "Delphi Rounds: Consensus and Stability"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Delphi Rounds: Consensus and Stability}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>")
library(contentvalidR)
```

## Two questions, not one

In a Delphi study, an expert panel rates the same items over several rounds.
Between rounds, experts see anonymous feedback on how the panel rated, and they
may revise their own ratings. Two different questions come out of that process:

- **Consensus**: do enough experts agree about an item now?
- **Stability**: are experts still changing their ratings?

A panel can agree while still shifting, and it can hold steady without
agreeing. Dajani, Sincoff and Talley (1979) argued that a level of agreement is
only worth interpreting once responses have stopped changing.
`delphi_validity()` therefore reports the two questions side by side and never
folds one into the other.

## Decide three things before the first round

Three choices belong in the study protocol, written down before any data
arrive, not in the analysis afterwards.

1. **What counts as agreeing.** On a 4-point relevance scale this is usually a
   rating of 3 or 4. In `delphi_validity()` it is `agree_cut`.
2. **The consensus threshold.** Diamond et al. (2014) reviewed 100 Delphi
   studies and recommend defining consensus before the study begins. The
   median threshold in the studies they reviewed was 75%. That figure describes
   common practice; it is not a validated cut-off. For this reason
   `consensus_threshold` has no default. Without one, results are reported
   descriptively.
3. **How stability will be measured.** The default is weighted kappa (below).
   Cohen (1968) points out that the weights are part of how agreement is
   defined, and so they must be chosen before the data are collected.

## The data

`delphi_validity()` takes long-format data: one row per expert, item, and
round. Here, ten experts rate six statements for relevance on a 1-4 scale over
three rounds. Two things happen that are common in real studies. Expert E10
leaves after round 2, and statement S1 reaches consensus in round 2 and is set
aside, so nobody rates it in round 3.

```{r data}
round1 <- cbind(
  S1 = c(4, 4, 3, 4, 3, 4, 2, 4, 3, 4),
  S2 = c(3, 4, 3, 2, 4, 3, 4, 3, 2, 4),
  S3 = c(2, 1, 2, 2, 1, 3, 2, 1, 2, 2),
  S4 = c(4, 2, 3, 1, 4, 2, 3, 1, 4, 2),
  S5 = c(2, 3, 2, 3, 2, 4, 3, 2, 3, 2),
  S6 = c(3, 2, 4, 1, 3, 2, 4, 1, 3, 2)
)
round2 <- cbind(
  S1 = c(4, 4, 4, 4, 3, 4, 3, 4, 3, 4),
  S2 = c(4, 4, 4, 4, 4, 4, 4, 3, 4, 4),
  S3 = c(2, 1, 2, 2, 1, 2, 2, 1, 2, 2),
  S4 = c(4, 2, 3, 1, 4, 1, 3, 2, 4, 2),
  S5 = c(3, 3, 2, 3, 3, 4, 3, 2, 3, 3),
  S6 = c(2, 3, 1, 4, 2, 3, 1, 4, 2, 3)
)
round3 <- cbind(  # E10 has left; S1 was set aside
  S2 = c(4, 4, 4, 4, 4, 4, 4, 4, 4),
  S3 = c(2, 1, 2, 2, 1, 2, 2, 1, 2),
  S4 = c(4, 1, 3, 1, 4, 1, 3, 2, 4),
  S5 = c(3, 3, 3, 3, 3, 4, 3, 3, 3),
  S6 = c(3, 2, 4, 1, 3, 2, 4, 1, 3)
)

to_long <- function(m, round) {
  data.frame(expert = paste0("E", seq_len(nrow(m))),
             item = rep(colnames(m), each = nrow(m)),
             round = round, rating = as.vector(m))
}
ratings <- rbind(to_long(round1, 1), to_long(round2, 2), to_long(round3, 3))
head(ratings)
```

If your data are already long, name the columns with `expert_col`,
`item_col`, `round_col`, and `rating_col`.

## Fitting the workflow

The threshold of 75% was fixed in this example's protocol before the study.
The seed makes the bootstrap intervals reproducible.

```{r fit}
fit <- delphi_validity(ratings, lo = 1, hi = 4, consensus_threshold = 0.75,
                       seed = 2026)
fit
```

The printout ends with a key to every column and status. The rest of this
article turns the key off, to keep the output short:

```{r key-off}
options(contentvalidR.show_key = FALSE)
```

## Reading consensus

`prop_agree` is the share of experts who rated the item at or above
`agree_cut` in its last round. On a relevance scale this is the item's I-CVI.
Three statements (S1, S2, and S5) reached the 75% threshold. S1's last round is
round 2, because it was set aside once it reached consensus.

Look at S3. Nobody rated it relevant, which looks like "no consensus". But the
panel does agree: every expert rated it below the cut. The interpretation says
so. Whether agreement in that direction counts as consensus to *exclude* an
item is for your protocol to define, not for the software to decide.

```{r interpretation}
shown <- fit$results[fit$results$item %in% c("S3", "S4"), ]
writeLines(strwrap(paste0(shown$item, ": ", shown$interpretation),
                   width = 72, exdent = 4))
```

## Reading stability

Read `prop_unchanged` first. It is the plainest measure: the share of experts
who gave exactly the same rating again. Then read the kappa trend beside it.
Kappa is weighted kappa between each expert's ratings in two consecutive rounds
(Holey et al., 2007). With the default quadratic weights, it equals the
intraclass correlation of the two rounds' ratings (Fleiss & Cohen, 1973). A
shift of the whole panel after feedback therefore counts as instability.

The six statements show the main patterns:

- **S3 and S4 are stable.** Almost every expert kept their rating, and kappa is
  high: between 0.83 and 1 in every pair of rounds. But S4 is stable *without*
  consensus: the panel is split (56%
  agreeing) and staying split. Another round is unlikely to change that.
  Dajani et al. (1979) treat a stable split as a result in its own right, to be
  reworded or reported rather than iterated away.
- **S2 converged to unanimity.** In round 3 every expert rated it 4. Of the
  experts who rated it in both rounds, 89% kept their rating, yet kappa is 0.
  When every rating in a round is the same, there is no room left above
  chance, so kappa is 0 no matter how many experts held steady. The output
  flags this and points you to `prop_unchanged`.
- **S5 formed consensus while kappa fell,** from 0.63 to 0.53. As ratings bunch
  near 3, kappa has less room to show agreement. This is the same pattern
  Holey et al. (2007) reported: their most-agreed statement had their lowest
  kappa. With nine or ten experts, the intervals run from about 0 to 1.
- **S6 is unstable.** No expert kept their rating, and kappa is negative
  because experts swapped positions.

These patterns are why the output gives kappa no verbal labels such as
"moderate" or "substantial". Landis and Koch (1977), who introduced those
labels, called their divisions clearly arbitrary. And a label would tend to
get *worse* as a Delphi succeeds, because convergence lowers kappa. Read kappa
as a trend, next to `prop_unchanged`.

The full table, with every pair of rounds, is in `details$stability`:

```{r stability}
st <- fit$details$stability
cols <- c("prop_unchanged", "value", "lower", "upper")
st[cols] <- round(st[cols], 2)
st[c("item", "from_round", "to_round", "n_paired", cols)]
```

`n_paired` counts the experts who rated the item in both rounds. It drops to 9
in the second pair because E10 left.

## Why the default looks at individual experts

S6 is the reason stability is measured expert by expert. Under Scheibe,
Skutsch and Schofer's (1975) 15% rule, which compares the two rounds' overall
rating distributions, S6 looks perfectly stable:

```{r percent-change}
pc <- delphi_validity(ratings, lo = 1, hi = 4, consensus_threshold = 0.75,
                      stability = "percent_change")
s6 <- pc$details$stability[pc$details$stability$item == "S6", ]
s6$value <- round(s6$value, 2)
s6[c("item", "from_round", "to_round", "prop_unchanged", "value", "stable")]
```

The distribution barely moved: the net change is 0% and then 11%. Yet no
expert kept their rating. Experts traded places, so the totals stayed the same.
Chaffin and Talley (1980) made exactly this point against group-level
stability tests. Every published method remains available through
`stability`, and each one prints its own limits:

| `stability` | What it measures | Reads as stable when | Main limitation |
|---|---|---|---|
| `"kappa"` (default) | Each expert's agreement with their own earlier rating, corrected for chance | Read as a trend; no cut-off | Falls as ratings converge |
| `"lambda"` | How well an earlier rating predicts the later one | Read as a trend | Predictability, not agreement; undefined if the later round is unanimous |
| `"chisq_individual"` | Whether later ratings depend on earlier ones | p < alpha | Association, not agreement; needs expected counts of 5 or more |
| `"chisq_group"` | Whether the two rounds' distributions differ | p >= alpha | Small panels look stable by default; misses experts swapping answers |
| `"percent_change"` | Net movement of the distribution | Change below 15% | No statistical basis; misses swapping; coarse with small panels |

The two chi-square tests read in opposite directions: one counts a significant
result as stable, the other a non-significant one. The printout always says
which is which.

## Relevance evidence round by round

Each round is also analyzed with `expert_validity()`, so the usual relevance
evidence (I-CVI, Aiken's V, and their intervals) is available round by round.
`compare_rounds()` reads those fits directly:

```{r compare}
do.call(compare_rounds, unname(fit$details$round_fits))
```

S1 shows as `Removed` in round 3 because it was set aside after reaching
consensus. S5 strengthened once experts converged.

## Reporting

`content_report()` gives a compact table for a manuscript:

```{r report}
content_report(fit)
```

Diamond et al. (2014) recommend that a Delphi report state the definition of
consensus and its threshold (decided before the study), the criteria for
stopping, and the criteria for dropping items. Also report:

- the number of rounds and experts in each round;
- the stability statistic, and the weights if it is kappa;
- the number of bootstrap resamples and the seed.

## Carrying a Delphi forward

When the panel is finished, `content_handoff()` packages the items that
survived, with the evidence behind each decision, for the empirical stage:

```{r handoff}
handoff <- content_handoff(fit, keep = "Supported")
handoff$item_evidence[c("item", "carried", "status", "n_judges", "round")]
```

`round` is the round each item settled in, so S1 carries round 2: it reached
consensus there and was set aside. Every other workflow writes one constant
into that column, because one fit is one round.

Each item's relevance evidence comes from its own last round, with intervals,
and the stability evidence travels beside it:

```{r handoff-statistics}
st <- handoff$item_statistics
unique(st$statistic)
```

Stability is carried as evidence, not as a reason to keep or drop an item, just
as it does not set an item's status above. See
`vignette("handoff-to-empirical-validation")` for what the object holds and how
a downstream package reads it.

## What this analysis does not establish

Consensus is agreement among these experts. It is not, by itself, evidence
that an item is valid. Feedback between rounds can also press experts toward
agreement they do not hold: Scheibe, Skutsch and Schofer (1975) observed that
participants who conformed most strongly were the least satisfied with the
process. A `No consensus` result is not an instruction to drop an item, and
`Consensus` is not an instruction to keep one. Read these results together
with the experts' comments, the construct definition, and the rest of the
content-validity evidence.

## References

Chaffin, W. W., & Talley, W. K. (1980). Individual stability in Delphi
studies. *Technological Forecasting and Social Change, 16*(1), 67-73.
https://doi.org/10.1016/0040-1625(80)90074-8

Cohen, J. (1968). Weighted kappa: Nominal scale agreement provision for scaled
disagreement or partial credit. *Psychological Bulletin, 70*(4), 213-220.
https://doi.org/10.1037/h0026256

Dajani, J. S., Sincoff, M. Z., & Talley, W. K. (1979). Stability and agreement
criteria for the termination of Delphi studies. *Technological Forecasting and
Social Change, 13*(1), 83-90. https://doi.org/10.1016/0040-1625(79)90007-6

Diamond, I. R., Grant, R. C., Feldman, B. M., Pencharz, P. B., Ling, S. C.,
Moore, A. M., & Wales, P. W. (2014). Defining consensus: A systematic review
recommends methodologic criteria for reporting of Delphi studies. *Journal of
Clinical Epidemiology, 67*(4), 401-409.
https://doi.org/10.1016/j.jclinepi.2013.12.002

Fleiss, J. L., & Cohen, J. (1973). The equivalence of weighted kappa and the
intraclass correlation coefficient as measures of reliability. *Educational and
Psychological Measurement, 33*(3), 613-619.
https://doi.org/10.1177/001316447303300309

Holey, E. A., Feeley, J. L., Dixon, J., & Whittaker, V. J. (2007). An
exploration of the use of simple statistics to measure consensus and stability
in Delphi studies. *BMC Medical Research Methodology, 7*, 52.
https://doi.org/10.1186/1471-2288-7-52

Landis, J. R., & Koch, G. G. (1977). The measurement of observer agreement for
categorical data. *Biometrics, 33*(1), 159-174. https://doi.org/10.2307/2529310

Scheibe, M., Skutsch, M., & Schofer, J. (1975). Experiments in Delphi
methodology. In H. A. Linstone & M. Turoff (Eds.), *The Delphi method:
Techniques and applications*. Addison-Wesley.
