---
title: "Design & Reporting Guide"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Design & Reporting Guide}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>")
```

```{r setup}
library(contentvalidR)
```

**Choosing a method**
- **Item-sort (Psa/Csv)**: fast, nominal assignments; good when raters map items → constructs.
- **Construct ratings (HTC/HTD + repeated-measures screening)**: the same judges rate each item against every focal/orbiting definition; useful for diagnosing definitional correspondence and overlap.
- **Classic indices** (CVR/IOC/Aiken’s V/CVI): expert-panel contexts; often complements Psa/Csv or ANOVA.

**Sample size tips**
- **Sorts**: choose the judge sample size in relation to the retention rule and the target-assignment probability you consider plausible. `sort_power()` provides exact planning probabilities; avoid treating a single rule of thumb as a universal minimum.
- **Rating studies**: the Hinkin-Tracey workflow is fully crossed within judges. Sample-size needs depend on the number of judges, number of construct definitions, and expected target-orbiting separation. Report the number of complete judge profiles used itemwise.
- **Panels**: Lawshe CVR tables are sensitive with small $N$; triangulate.

**Recommended reporting**
1. **Design** (instructions, coder training, scale anchors).
2. **Decision rules** (alpha, thresholds, tie-breaks).
3. **Primary results** (tables below).
4. **Diagnostics** (signal detection vs. later CFA/IRT; reproducibility across pilots).
5. **Data & code** (reproducible script, package version).

**Reproducible toy data used below**
```{r}
# --- Sort-based toy data (4 items x 12 raters) ---
toy_sort <- data.frame(
  item = rep(paste0("I", 1:4), each = 12),
  rater = rep(1:12, 4),
  target_construct   = rep(c("A","A","B","B"), each = 12),
  assigned_construct = c(
    sample(c("A","B"), 12, TRUE, c(.80,.20)), # I1
    sample(c("A","B"), 12, TRUE, c(.65,.35)), # I2
    sample(c("A","B"), 12, TRUE, c(.70,.30)), # I3
    sample(c("A","B"), 12, TRUE, c(.45,.55))  # I4 (likely weak)
  )
)
psa <- compute_psa(toy_sort)
csv <- compute_csv(toy_sort)
csv$decision <- vapply(seq_len(nrow(csv)), function(i) {
  csv_binom_test(n_c = csv$n_target[i], N = csv$n[i])$decision
}, character(1))

# --- Fully crossed construct-rating toy data ---
set.seed(2)
toy_ratings <- expand.grid(
  item = c("I1", "I2", "I3"),
  rater = 1:16,
  construct = c("A", "B", "C")
)
toy_ratings$target_construct <- ifelse(toy_ratings$item == "I3", "B", "A")
toy_ratings$rating <- ifelse(
  toy_ratings$construct == toy_ratings$target_construct,
  pmin(5, pmax(1, round(rnorm(nrow(toy_ratings), 4.4, .6)))),
  pmin(5, pmax(1, round(rnorm(nrow(toy_ratings), 2.2, .7))))
)
rating_out <- rating_validity(toy_ratings, scale_min = 1, scale_max = 5)
```

**Table templates**

**Sort-Based (per item)**
```{r}
csv$psa <- compute_psa(toy_sort)$psa
csv$binom_p <- vapply(seq_len(nrow(csv)), function(i) {
  csv_binom_test(csv$n_target[i], csv$n[i])$p.value
}, numeric(1))
csv_out <- csv[, c("item","n","n_target","n_other_max","psa","csv","binom_p","decision")]
csv_out
```

| item | N   | n<sub>target</sub> | n<sub>other_max</sub> | Psa  | Csv  | binom p | decision    |
|:-----|:----|:-------------------|:--------------------|:-----|:-----|:--------|:------------|
| I1   | 20  | 16                 | 3                   | .80  | .65  | .002    | significant |
| ...  | ... | ...                | ...                 | ...  | ...  | ...     | ...         |

**Construct-rating evidence (per item)**
```{r}
rating_out$results
```

Report HTC/HTD, the repeated-measures omnibus result, the planned target-versus-orbiting contrast result, the strongest competitor, and the package's Retain/Review screening flag. At the target-scale level, report mean HTC/HTD and the applicable Colquitt normative band.

**Classic indices (panel)**
```{r}
# Aiken's V
R <- matrix(sample(1:5, 5*6, replace = TRUE), nrow = 5,
            dimnames = list(NULL, paste0("Item", 1:6)))
aikens_v(R, lo = 1, hi = 5)

# CVR
cvr(essential = c(8,10,5), N = 12)

# CVI
M <- matrix(sample(0:1, 6*5, replace = TRUE, prob = c(.3,.7)), nrow = 6,
            dimnames = list(NULL, paste0("Item", 1:5)))
cvi(M)

# IOC
ioc_df <- data.frame(
  item = rep(paste0("I",1:2), each = 9),
  judge = rep(1:3, times = 6),
  objective = rep(rep(LETTERS[1:3], each = 3), times = 2),
  score = sample(c(-1,0,1), 18, replace = TRUE)
)
ioc(ioc_df)
```

| item  | Aiken's V | CVR  | I-CVI | S-CVI/Ave | IOC mean |
|:------|:----------|:-----|:------|:----------|:---------|
| Item1 | 0.86      | 0.67 | .80   | .88       | .72      |
| ...   | ...       | ...  | ...   | ...       | ...      |

**Diagnostics & reproducibility**
```{r}
truth <- c(TRUE, TRUE, TRUE, FALSE)  # pretend CFA-retained
signal_detection(csv$decision == "significant", truth)

csv2_sig <- sample(c(TRUE, FALSE), nrow(csv), replace = TRUE)
reproducibility_phi(csv$decision == "significant", csv2_sig)
```

**Power quick-checks**
```{r}
sort_power(N = c(20, 30), true_p = c(.65, .75))
```

**Good practices**
-Pre-register criteria where possible.
- Keep README + vignette synchronized; export data/code to a replication repo.
- For multi-item scales, consider multiplicity adjustments only if you formalize per-item hypothesis testing; otherwise treat as screening + triangulation.
