---
title: "Design & Reporting Guide"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Design & Reporting Guide}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>")
```

```{r setup}
library(contentvalidR)
```

**Choosing a method**
- **Item-sort (Psa/Csv)**: fast, nominal assignments; good when raters map items → constructs.
- **Ratings + ANOVA**: Likert ratings across all constructs; stronger convergent/discriminant evidence.
- **Classic indices** (CVR/IOC/Aiken’s V/CVI): expert-panel contexts; often complements Psa/Csv or ANOVA.

**Sample size tips**
- **Sorts**: $N ≥ 15$ raters gives workable binomial power; $20–40$ ideal.
- **ANOVA**: balanced $n$ per construct (e.g., $n ≥ 10$).
- **Panels**: Lawshe CVR tables are sensitive with small $N$; triangulate.

**Recommended reporting**
1. **Design** (instructions, coder training, scale anchors).
2. **Decision rules** (alpha, thresholds, tie-breaks).
3. **Primary results** (tables below).
4. **Diagnostics** (signal detection vs. later CFA/IRT; reproducibility across pilots).
5. **Data & code** (reproducible script, package version).

**Reproducible toy data used below**
```{r}
# --- Sort-based toy data (4 items x 12 raters) ---
toy_sort <- data.frame(
  item = rep(paste0("I", 1:4), each = 12),
  rater = rep(1:12, 4),
  target_construct   = rep(c("A","A","B","B"), each = 12),
  assigned_construct = c(
    sample(c("A","B"), 12, TRUE, c(.80,.20)), # I1
    sample(c("A","B"), 12, TRUE, c(.65,.35)), # I2
    sample(c("A","B"), 12, TRUE, c(.70,.30)), # I3
    sample(c("A","B"), 12, TRUE, c(.45,.55))  # I4 (likely weak)
  )
)
psa <- compute_psa(toy_sort)
csv <- compute_csv(toy_sort)
csv$decision <- vapply(seq_len(nrow(csv)), function(i) {
  csv_binom_test(n_c = csv$n_target[i], N = csv$n[i])$decision
}, character(1))

# --- Rating-based toy data for ANOVA (items rated on all constructs) ---
set.seed(2)
toy_ratings <- data.frame(
  item = rep(paste0("I",1:4), each = 30),
  rater = rep(1:30, times = 4),
  construct = rep(rep(LETTERS[1:3], each = 10), times = 4),
  rating = c(
    rnorm(10, 4.3, 0.4), rnorm(20, 2.6, 0.6),   # I1: A high
    rnorm(10, 4.0, 0.5), rnorm(20, 2.7, 0.6),   # I2: A high
    rnorm(10, 2.6, 0.6), rnorm(10, 4.1, 0.5), rnorm(10, 2.6, 0.6), # I3: B high
    rnorm(20, 2.6, 0.6), rnorm(10, 4.0, 0.5)    # I4: C high
  )
)
target_map <- c(I1="A", I2="A", I3="B", I4="C")
anova_out <- anova_content(toy_ratings, target_map = target_map)
```

**Table templates**
**Sort-Based (per item)**
```{r}
csv$psa <- compute_psa(toy_sort)$psa
csv$binom_p <- vapply(seq_len(nrow(csv)), function(i) {
  csv_binom_test(csv$n_target[i], csv$n[i])$p.value
}, numeric(1))
csv_out <- csv[, c("item","n","n_target","n_other_max","psa","csv","binom_p","decision")]
csv_out
```

| item | N   | n<sub>target</sub> | n<sub>other_max</sub> | Psa  | Csv  | binom p | decision    |
|:-----|:----|:-------------------|:--------------------|:-----|:-----|:--------|:------------|
| I1   | 20  | 16                 | 3                   | .80  | .65  | .002    | significant |
| ...  | ... | ...                | ...                 | ...  | ...  | ...     | ...         |

**Rating-based ANOVA (per item)**
```{r}
anova_out
```

| item | target | F(df1,df2) | p      | posthoc_pass |
|:-----|:-------|:-----------|:-------|:-------------|
| I1   | A      | 12.4 (2,27)  | .0001  | TRUE         |
| ...  | ...    | ...        | ...    | ...          |

**Classic indices (panel)**
```{r}
# Aiken's V
R <- matrix(sample(1:5, 5*6, replace = TRUE), nrow = 5,
            dimnames = list(NULL, paste0("Item", 1:6)))
aikens_v(R, lo = 1, hi = 5)

# CVR
cvr(essential = c(8,10,5), N = 12)

# CVI
M <- matrix(sample(0:1, 6*5, replace = TRUE, prob = c(.3,.7)), nrow = 6,
            dimnames = list(NULL, paste0("Item", 1:5)))
cvi(M)

# IOC
ioc_df <- data.frame(
  item = rep(paste0("I",1:2), each = 9),
  judge = rep(1:3, times = 6),
  objective = rep(rep(LETTERS[1:3], each = 3), times = 2),
  score = sample(c(-1,0,1), 18, replace = TRUE)
)
ioc(ioc_df)
```

| item  | Aiken's V | CVR  | I-CVI | S-CVI/Ave | IOC mean |
|:------|:----------|:-----|:------|:----------|:---------|
| Item1 | 0.86      | 0.67 | .80   | .88       | .72      |
| ...   | ...       | ...  | ...   | ...       | ...      |

**Diagnostics & reproducibility**
```{r}
truth <- c(TRUE, TRUE, TRUE, FALSE)  # pretend CFA-retained
signal_detection(csv$decision == "significant", truth)

csv2_sig <- sample(c(TRUE, FALSE), nrow(csv), replace = TRUE)
reproducibility_phi(csv$decision == "significant", csv2_sig)
```

**Power quick-checks**
```{r}
simulate_csv_power(N = 20, true_p = 0.65, reps = 200)
simulate_anova_power(n_raters = 20, mean_diff = 0.5, sd = 1, k_constructs = 3, reps = 100)
```

**Good practices**
-Pre-register criteria where possible.
- Keep README + vignette synchronized; export data/code to a replication repo.
- For multi-item scales, consider multiplicity adjustments only if you formalize per-item hypothesis testing; otherwise treat as screening + triangulation.
