---
title: "Design & Reporting Guide"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Design & Reporting Guide}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>")
library(contentvalidR)
read_example <- function(name) {
  utils::read.csv(
    system.file("extdata", name, package = "contentvalidR"),
    stringsAsFactors = FALSE
  )
}
```

## Choosing a method

Use the design that matches the question posed to judges rather than selecting an
index after data collection.

- **Item sort:** judges assign each item to the construct definition that best
  represents it. Use `sort_validity()` for Psa/Csv, exact target-count screening,
  strongest-competitor diagnostics, and scale-level Colquitt norms.
- **Construct rating:** the same judges rate each item against every focal and
  orbiting definition. Use `rating_validity()` for HTC/HTD, repeated-measures
  inference, planned target-versus-orbiting contrasts, and scale-level norms.
- **Expert panel:** experts answer a relevance, essentiality, or congruence
  question. Use `expert_validity()` with the corresponding explicit `mode`.
- **Delphi rounds:** the same panel rates the items over several rounds, with
  feedback in between. Use `delphi_validity()` for consensus against a
  threshold fixed before the study, and stability between rounds.

The designs can complement one another during scale development, but their
statistics are not interchangeable. Two further questions apply to any of them:
whether the conclusions depend on the particular judges who served
(`judge_validity()`), and whether the items cover the blueprint of the domain
(`domain_validity()`).

## Sample-size planning

For item sorts, plan judge N in relation to the exact retention rule and a
plausible true target-assignment probability. `sort_power()` provides exact
planning probabilities; avoid a universal judge-count rule of thumb.

For construct-rating studies, power depends on the number of judges, number of
construct definitions, within-judge target-orbiting separation, and missing
profiles. Report the effective complete-judge N itemwise.

For expert panels, panel-size sensitivity is part of the statistic. CVR exact
critical counts and common CVI review guidelines therefore need to be interpreted
with the actual effective N, not a nominal panel size that ignores missingness.

## Recommended reporting sequence

1. **Design:** judge population, recruitment, construct definitions, item pool,
   instructions, and response format.
2. **A priori rules:** alpha, relevance threshold, target mapping, multiplicity
   adjustment if used, and any planned norms.
3. **Item-level evidence:** correspondence, distinctiveness, exact/paired
   inference, strongest competitor, and review status as appropriate.
4. **Scale-level evidence:** target-scale averages or S-CVI summaries where the
   method defines them.
5. **Substantive decisions:** revisions, removals, retained domain-coverage
   items, and how qualitative comments informed those decisions.
6. **Reproducibility:** package version, analysis settings, anonymized data when
   permitted, and the script used to reproduce tables/figures.

The `reporting-examples` vignette expands this sequence into reusable methods and
results scaffolds. Treat those examples as reporting patterns rather than fixed
language that must be copied verbatim.

## Deterministic bundled examples

```{r example-data}
sort_dat <- read_example("sort_example.csv")
rating_dat <- read_example("rating_example.csv")
expert_rel <- read_example("expert_relevance_example.csv")

sort_fit <- sort_validity(sort_dat)
rating_fit <- rating_validity(rating_dat, scale_min = 1, scale_max = 5)
expert_fit <- expert_validity(
  as.matrix(expert_rel[setdiff(names(expert_rel), "expert")]),
  mode = "relevance", lo = 1, hi = 4
)
```

These files are synthetic and generated by `data-raw/build-example-data.R` in the
source repository. They deliberately include both supported and review-worthy
items so documentation exercises realistic output paths without depending on
random-number generation.

## Table templates

`content_report()` turns a fit into a manuscript table, with the columns a
reader needs and numbers in APA form; `format = "markdown"` gives the same
table to paste. The unrounded values stay in each fit's `results`.

### Item sort

```{r sort-table}
content_report(sort_fit)
```

At the target-scale level, report mean Psa/Csv and the benchmark set actually
used. Do not convert Colquitt's scale-level norms into individual-item cutoffs.

### Construct rating

```{r rating-table}
content_report(rating_fit)
```

Report the repeated-measures design and target-versus-orbiting contrasts. For a
review item, naming the strongest competitor, which `rating_fit$results` holds
as `strongest_competitor`, is often more informative than a standalone p
value.

### Expert relevance

```{r expert-table}
content_report(expert_fit)
expert_fit$scale_summary[c("S_CVI_Ave", "S_CVI_UA", "agreement")]
```

Essentiality and congruence require different expert tasks. Do not place CVR,
CVI, Aiken V, and IOC in one generic threshold table as if they answer the same
question.

## Diagnostics and later validation

`signal_detection()` and `reproducibility_phi()` remain auxiliary helpers. They
can be useful when researchers later compare pretest decisions with CFA/IRT
retention or an independent replication pretest, but they are not required parts
of the flagship workflows.

```{r diagnostics}
pretest_supported <- sort_fit$results$status == "Supported"
later_retained <- c(TRUE, TRUE, TRUE, FALSE, TRUE, FALSE)
signal_detection(pretest_supported, later_retained)

replication_supported <- c(TRUE, TRUE, TRUE, FALSE, TRUE, TRUE)
reproducibility_phi(pretest_supported, replication_supported)
```

With six items every expected count is below 5, so both helpers report
Fisher's exact *p* and show the chi-square approximation beside it for
reference. With enough items for the approximation they report the chi-square
test, with its degrees of freedom and *N*.

## Planning how many judges to recruit

`sort_power()` and `expert_power()` answer the question exactly for one
design, from the decision rule the analysis will apply, rather than from a
rule of thumb. `gtheory_content()` estimates it from a panel's ratings.

For an item sort, `sort_power()` gives the probability that an item reaches the
exact retention count, for each panel size and each rate at which you expect
judges to choose the target:

```{r power}
sort_power(N = c(20, 30, 40), true_p = c(.60, .70, .80))
```

For an expert panel, `expert_power()` gives the probability that an item clears
the I-CVI criterion, for each panel size and an assumed probability that one
expert endorses it:

```{r expert-planning}
expert_power(n_experts = 3:8, prob = 0.9)
```

Note the step. Lynn's (1986) criterion requires every expert to agree on panels
of up to five, and allows one disagreement from six, so a fourth or fifth
expert **lowers** the probability of clearing while a sixth raises it sharply.
That is a property of the guideline, not of the items, and the package reports
it rather than smoothing it away.

When a panel has already rated the items, generalizability theory (Brennan,
2001; Crocker et al., 1988) estimates how many judges the same design would
need to reach a given dependability:

```{r gtheory}
judge_ratings <- rbind(
  c(4, 4, 4, 3, 2, 2), c(4, 4, 3, 4, 2, 1), c(4, 3, 4, 4, 1, 2),
  c(3, 4, 4, 4, 2, 2), c(4, 4, 4, 4, 2, 1), c(4, 3, 4, 3, 1, 2),
  c(4, 4, 3, 4, 2, 2), c(2, 2, 2, 2, 1, 1)
)
dimnames(judge_ratings) <- list(paste0("Judge", 1:8), paste0("Item", 1:6))
gtheory_content(judge_ratings)
```

The dependability coefficient (Phi) concerns the absolute level of the ratings
and is lowered by differences in judge severity, which is usually what a
content-validity decision rests on. The status beside it compares Phi with
`phi_cut` (.80 by default), a convention of this package, not a published
standard. Here every target is reached with five judges or fewer, because the
items account for most of the variance. Had the judges barely distinguished the
items, a target no realistic panel reaches would print as "unreachable" (`NA`
in the data). The projected panel sizes come from one panel's variance
components and carry no interval, so read them as planning figures.

Across pretest rounds, `compare_rounds()` sets each item's status side by side
and checks whether the analysis settings changed between rounds, so a relaxed
criterion cannot read as an improved item.

## Good practices

Pre-register quantitative screening criteria when feasible, preserve qualitative
judge feedback, report effective N after missingness, and archive the construct
definitions and item wording used in the pretest. A content-validation statistic
is evidence from a designed judgment task; it is not a substitute for defining
and sampling the construct domain.

## References

Brennan, R. L. (2001). *Generalizability theory*. Springer.
https://doi.org/10.1007/978-1-4757-3456-0

Crocker, L., Llabre, M., & Miller, M. D. (1988). The generalizability of content
validity ratings. *Journal of Educational Measurement, 25*(4), 287–299.
https://doi.org/10.1111/j.1745-3984.1988.tb00309.x

Lynn, M. R. (1986). Determination and quantification of content validity.
*Nursing Research, 35*(6), 382–385.
https://doi.org/10.1097/00006199-198611000-00017
