---
title: "Expert-Panel Content Validation: Relevance, Essentiality, and Congruence"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Expert-Panel Content Validation: Relevance, Essentiality, and Congruence}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>")
library(contentvalidR)
```

## Why expert-panel methods need separate modes

Expert-panel content validation is not one statistical task. A relevance rating,
an essential/not-essential judgment, and an item-objective congruence judgment
ask experts different questions and therefore support different indices.
`expert_validity()` uses an explicit `mode` so those designs are not treated as
interchangeable.

The three modes are:

- **relevance**: Aiken's V plus CVI and modified kappa, with a panel-level
  agreement coefficient;
- **essentiality**: Lawshe's CVR with exact binomial inference; and
- **congruence**: the index of item-objective congruence (IOC) of Rovinelli
  and Hambleton (1977).

All three are quantitative complements to qualitative expert comments, construct
coverage, comprehensibility review, and other parts of the content-validity
argument.

## Relevance: Aiken V, score intervals, and CVI

Suppose six experts rate item relevance from 1 (not relevant) to 4 (highly
relevant):

```{r relevance}
R <- matrix(
  c(4,4,4,4,4,4,
    4,4,4,3,4,4,
    4,3,4,4,3,4,
    3,3,4,3,2,3),
  nrow = 6,
  dimnames = list(NULL, paste0("Item", 1:4))
)

fit <- expert_validity(R, mode = "relevance", lo = 1, hi = 4, seed = 1)
fit
summary(fit)
```

Aiken's V rescales the bounded expert ratings to the 0-1 interval. The default
confidence interval is the score interval proposed by Penfield and Giacobbi
(2004), rather than a simulation-dependent bootstrap interval. Bootstrap
intervals remain available through the low-level function:

```{r aiken-bootstrap}
aikens_v(R, lo = 1, hi = 4, ci = "bootstrap", B = 200, seed = 1)
```

For CVI, the workflow dichotomizes ratings at `relevance_cut`. On a 1-4 scale
the default is 3, so ratings of 3 or 4 count as relevant. Declare a different
threshold if the study protocol used one.

The workflow reports Lynn's (1986) panel-size criterion for the I-CVI as a
**review aid**: with three to five experts all must agree, from six one may
disagree, and from nine two may (7 of 9, the .78 usually quoted). Lynn's table
stops at ten experts; beyond that the package holds her lowest proportion, 7 of
9, an extension of her rule that the printout and the handoff label as this
package's own. Polit and Beck (2006) restate her rule as no lower than .78 for
six or more experts, though their own recommendation covers six to ten. The criterion is not presented as universal proof that an item is or is not
content valid. Modified kappa provides
a chance-corrected complement to I-CVI.

At the scale level, S-CVI/Ave and S-CVI/UA are reported together. S-CVI/Ave is
generally less brittle than universal agreement, but both should be interpreted
alongside the distribution of item-level evidence.

## Panel-level agreement

I-CVI and modified kappa describe one item at a time. Relevance mode also
reports how consistently the panel rated the whole item set, as one coefficient
with a bootstrap interval:

```{r agreement}
fit$scale_summary[, c("agreement", "agreement_low", "agreement_high")]
fit$details$agreement
```

The default coefficient is Krippendorff's alpha. It accepts any number of
experts and missing ratings, and Zapf et al. (2016) recommend it when ratings
are ordinal or incomplete, which describes most expert panels. It is a general
reliability coefficient (Hayes & Krippendorff, 2007) rather than one developed
for content validity; no publication applying it specifically to
content-validity panels was found.

Choose the measurement level that matches the rating scale. Relevance ratings
are treated as ordinal by default. `agreement_level = "interval"` treats the
distances between scale points as equal, and `"nominal"` treats every
disagreement as equally serious:

```{r agreement-level}
expert_validity(R, mode = "relevance", lo = 1, hi = 4,
                agreement_level = "interval", agreement_B = 0)$scale_summary$agreement
```

### Why alpha can be low when experts agree

Alpha compares the disagreement within items with the disagreement expected if
the same ratings were scattered across items at random. When a panel rates
nearly every item 4, very little disagreement is expected by chance, so a few 3s
pull alpha down even though most rating pairs are identical. Feinstein and
Cicchetti (1990) described the same pattern for kappa. The output reports the
share of identical rating pairs next to alpha so the two can be read together.
A low alpha alongside a high share of identical pairs is not by itself evidence
of a poor panel.

### Gwet's AC1

Gwet's (2008) AC1 was designed to stay high in that situation, and it is
available with `agreement = "ac1"`. It is never the default. Vach and Gerke
(2023) show that AC1 rises as ratings concentrate in one category even when
agreement does not change, and that it can be above zero when experts rate
independently. Its output always repeats that critique. In relevance mode, AC1
is computed on the relevant/not-relevant decision at `relevance_cut`:

```{r agreement-ac1}
ac1_fit <- expert_validity(R, mode = "relevance", lo = 1, hi = 4,
                           agreement = "ac1", agreement_B = 0)
ac1_fit$details$agreement
```

### The interval

The interval resamples items with all of their ratings intact, the procedure
Zapf et al. (2016) evaluated; they found that Krippendorff's original bootstrap,
which ignores dependence between raters, reached only about 60% coverage. The
interval varies slightly between runs, so set `seed` to make it reproducible, or
set `agreement_B = 0` to skip it. `panel_agreement()` runs the same analysis on
any rater-by-item matrix.

## Essentiality: Lawshe CVR with exact critical values

Lawshe's task asks experts whether an item is essential. With twelve experts:

```{r essentiality}
expert_validity(c(10, 8, 6), mode = "essentiality", N = 12)
```

`cvr()` derives the smallest essential count whose one-sided binomial upper-tail
probability is no greater than `alpha`. This makes the panel-size dependency
explicit and follows the exact-probability logic revisited by Ayre and Scally
(2014).

Judge-by-item binary data can be supplied directly:

```{r essential-matrix}
E <- cbind(
  Item1 = c(1,1,1,1,1,1,1,1),
  Item2 = c(1,1,1,1,1,0,0,0)
)
expert_validity(E, mode = "essentiality")
```

A failure to clear the exact criterion is labeled `Review`, not automatic
deletion. Expert rationales and domain coverage matter when deciding whether an
item should be rewritten, retained for breadth, or removed.

With four or fewer experts no count can clear the criterion at the default
alpha: all four calling an item essential gives *p* = .0625. Such an item is
labeled `Insufficient panel`, not `Review`, because the panel is too small to
decide, whatever the experts said.

### Lawshe's table and its recalculation, for comparison

Lawshe (1975) published a table of minimum CVR values, and Wilson et al.
(2012) recalculated it. `legacy = TRUE` prints both beside the exact
decision, without changing it:

```{r essentiality-legacy}
old <- options(contentvalidR.show_key = FALSE)
expert_validity(c(9, 8, 7), mode = "essentiality", N = 9, legacy = TRUE)
options(old)
```

Nine experts show why the table was questioned. Its minimum rises steadily
from 40 experts down to nine, then drops from .78 at nine to .75 at eight,
which Wilson et al. (2012) treated as an anomaly. Ayre and Scally (2014)
showed it is not one. With so few experts only a few CVR values are possible:
.75 is 7 of 8, and .78 is 8 of 9 (.778) printed to two decimals, the same
counts the exact test requires.

In relevance mode, `legacy = TRUE` adds Fleiss' (1971) kappa on the
relevant/not-relevant decision and the Polit and Beck (2006) benchmarks for
S-CVI/Ave and S-CVI/UA. It also adds the content validity coefficient (Ccv)
of Hernández-Nieto (2002), which is widely cited, beside Aiken's V. Read it as a
cautionary example rather than evidence. Ccv is the mean rating divided by the
scale maximum, minus a "chance" term, (1/J)^J, that depends only on the number
of judges and is .00032 for five of them. So on a scale starting at 0 it is
Aiken's V, and on a scale starting at 1 it cannot fall below one over the
scale maximum. It uses only the mean, so despite the book's claim it cannot
reflect agreement: in the book's own Table 7, ratings of 1, 3, 4, 5, 2 and of
3, 3, 3, 3, 3 both score .60. Its .80 cutoff is stated without derivation, and
the book's own examples do not keep to it. `?expert_validity` sets out each of
these with page references.

## Congruence: item-objective alignment

Each expert rates each item against every objective: +1 if the item clearly
measures it, 0 if unclear, -1 if it clearly does not. The index of Rovinelli and
Hambleton (1977) for an objective is half the gap between the experts' mean rating
on that objective and their mean rating on the item's other objectives, so it
reaches 1 only when every expert gives +1 to the objective and -1 to every
other. Turner and Carlson (2003) extend the index to items written for more
than one objective, which this package does not do. A target mapping says
which objective each item was written for:

```{r congruence}
d <- expand.grid(
  item = c("I1", "I2"),
  judge = 1:4,
  objective = c("A", "B")
)
d$target_objective <- ifelse(d$item == "I1", "A", "B")
d$score <- ifelse(d$objective == d$target_objective, 1, -1)

expert_validity(d, mode = "congruence")
```

An item is `"Congruent"` when its index for the target objective reaches
`ioc_cut`, by default the .70 Rovinelli and Hambleton applied. The printout
also shows the experts' mean rating on the target and on the closest other
objective, and the margin between those means. Applied work often reports
the mean rating on the target as "the IOC"; it is an ingredient of the index,
not the index, because it reaches 1 whatever the experts say about the other
objectives. The margin is a description beside the index, not part of its
criterion. Without a target mapping, each item's index on every objective is
described, with no decision.

## Missing ratings

Missing data are never silently ignored by default. Set `na.rm = TRUE` only when
itemwise/cellwise deletion matches the study protocol. Effective expert counts
and missing counts are then reported so downstream interpretation uses the
actual panel size.

## Plotting expert evidence

Each mode uses a plot matched to the expert task rather than forcing unlike
indices into one generic chart.

```{r expert-plots, fig.width=7, fig.height=4, fig.alt=c("Expert relevance plot: for each item, Aiken's V (filled) and I-CVI (open), each with its 95% interval, and a dashed line at the I-CVI criterion for the panel size.", "Essentiality plot: each item's observed CVR (filled) joined to the CVR the exact test needs (open), with a dotted line at zero.", "Congruence plot: each item's index of item-objective congruence for its intended objective (filled), with the experts' mean rating on the intended objective (triangle) and on the closest other objective (open circle) in gray, a dotted line at zero, and a dashed line at the .70 criterion.")}
plot(expert_validity(R, mode = "relevance", lo = 1, hi = 4))
plot(expert_validity(c(10, 8, 6), mode = "essentiality", N = 12))
plot(expert_validity(d, mode = "congruence"))
```

Relevance mode displays Aiken's V with its score interval and overlays I-CVI as
a separate marker. Essentiality mode displays observed CVR against the exact
panel-specific critical CVR, and marks with a cross an item whose panel was too
small to test. Congruence mode places each item's index against the
criterion, with the two mean ratings behind it in gray. These displays are
diagnostic summaries; they do not create new validity thresholds.

An index reduces each item's ratings to one number, and the relevance plot
shows that number. The distribution view shows the ratings behind it, as
diverging stacked bars (Heiberger & Robbins, 2014). Each bar splits at the
relevance cut: ratings below it extend left and ratings at or above it extend
right, so the right-hand length is the I-CVI, read against the dashed criterion
line.

```{r expert-distribution, fig.width=7, fig.height=4, fig.alt="Rating distributions for four items rated by six experts, as bars split at the relevance cut of 3, with a dashed line at the I-CVI criterion of .83. Items 1 to 3 are rated relevant by all six experts, Item 1 entirely with 4s; Item 4 is rated relevant by five of six, mostly with 3s."}
plot(fit, type = "distribution",
     labels = c("Not relevant", "Somewhat", "Quite", "Highly relevant"))
```

Item1 and Item3 have the same I-CVI, but every expert gave Item1 a 4, while
two gave Item3 a 3. The index cannot tell them apart; the bars can, and so can
Aiken's V, which uses the whole scale. The figure is gray by default, as an APA
figure is printed. For slides or a poster, `apa = FALSE` draws it in color:
brown for ratings below the cut and teal for ratings at or above it. The
symbol beside each bar still carries the decision, so no reading depends on
color.

## Reporting

A concise methods/results description should identify:

1. who the experts were and why they were qualified;
2. the exact task and response scale;
3. the index and inference/CI procedure used, and for relevance ratings the
   agreement coefficient and its measurement level;
4. the panel size, including item-specific missingness;
5. quantitative item and scale evidence; and
6. how expert comments, construct coverage, and comprehensibility informed the
   final item decisions.

A content-validity coefficient is evidence about a defined expert task. It is
not, by itself, a complete validity argument.

## References

Aiken, L. R. (1980). Content validity and reliability of single items or
questionnaires. *Educational and Psychological Measurement, 40*(4), 955–959. https://doi.org/10.1177/001316448004000419

Ayre, C., & Scally, A. J. (2014). Critical values for Lawshe's content validity
ratio: Revisiting the original methods of calculation. *Measurement and
Evaluation in Counseling and Development, 47*(1), 79–86.
https://doi.org/10.1177/0748175613513808

Feinstein, A. R., & Cicchetti, D. V. (1990). High agreement but low kappa: I.
The problems of two paradoxes. *Journal of Clinical Epidemiology, 43*(6),
543–549. https://doi.org/10.1016/0895-4356(90)90158-L

Fleiss, J. L. (1971). Measuring nominal scale agreement among many raters.
*Psychological Bulletin, 76*(5), 378–382. https://doi.org/10.1037/h0031619

Gwet, K. L. (2008). Computing inter-rater reliability and its variance in the
presence of high agreement. *British Journal of Mathematical and Statistical
Psychology, 61*(1), 29–48. https://doi.org/10.1348/000711006X126600

Hayes, A. F., & Krippendorff, K. (2007). Answering the call for a standard
reliability measure for coding data. *Communication Methods and Measures, 1*(1),
77–89. https://doi.org/10.1080/19312450709336664

Heiberger, R. M., & Robbins, N. B. (2014). Design of diverging stacked bar
charts for Likert scales and other applications. *Journal of Statistical
Software, 57*(5), 1–32. https://doi.org/10.18637/jss.v057.i05

Hernández-Nieto, R. (2002). *Contributions to statistical analysis: The
coefficients of proportional variance, content validity and kappa*.
BookSurge.

Krippendorff, K. (2011). *Computing Krippendorff's alpha-reliability*. Annenberg
School for Communication, University of Pennsylvania.
https://www.asc.upenn.edu/sites/default/files/2021-03/Computing%20Krippendorff%27s%20Alpha-Reliability.pdf

Lawshe, C. H. (1975). A quantitative approach to content validity. *Personnel
Psychology, 28*(4), 563–575. https://doi.org/10.1111/j.1744-6570.1975.tb01393.x

Lynn, M. R. (1986). Determination and quantification of content validity.
*Nursing Research, 35*(6), 382–385.
https://doi.org/10.1097/00006199-198611000-00017

Penfield, R. D., & Giacobbi, P. R., Jr. (2004). Applying a score confidence
interval to Aiken's item content-relevance index. *Measurement in Physical
Education and Exercise Science, 8*(4), 213–225.
https://doi.org/10.1207/S15327841MPEE0804_3

Polit, D. F., & Beck, C. T. (2006). The content validity index: Are you sure
you know what's being reported? Critique and recommendations. *Research in
Nursing & Health, 29*(5), 489–497. https://doi.org/10.1002/nur.20147

Polit, D. F., Beck, C. T., & Owen, S. V. (2007). Is the CVI an acceptable
indicator of content validity? Appraisal and recommendations. *Research in
Nursing & Health, 30*(4), 459–467. https://doi.org/10.1002/nur.20199

Rovinelli, R. J., & Hambleton, R. K. (1977). On the use of content specialists
in the assessment of criterion-referenced test item validity. *Dutch Journal of
Educational Research, 2*, 49–60.

Turner, R. C., & Carlson, L. (2003). Indexes of item-objective congruence for
multidimensional items. *International Journal of Testing, 3*(2), 163–171.
https://doi.org/10.1207/S15327574IJT0302_5

Vach, W., & Gerke, O. (2023). Gwet's AC1 is not a substitute for Cohen's kappa:
A comparison of basic properties. *MethodsX, 10*, 102212.
https://doi.org/10.1016/j.mex.2023.102212

Wilson, F. R., Pan, W., & Schumsky, D. A. (2012). Recalculation of the critical
values for Lawshe's content validity ratio. *Measurement and Evaluation in
Counseling and Development, 45*(3), 197–210.
https://doi.org/10.1177/0748175612440286

Zapf, A., Castell, S., Morawietz, L., & Karch, A. (2016). Measuring inter-rater
reliability for nominal data: Which coefficients and confidence intervals are
appropriate? *BMC Medical Research Methodology, 16*, 93.
https://doi.org/10.1186/s12874-016-0200-9
