---
title: "Expert-Panel Content Validation: Relevance, Essentiality, and Congruence"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Expert-Panel Content Validation: Relevance, Essentiality, and Congruence}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>")
library(contentvalidR)
```

## Why expert-panel methods need separate modes

Expert-panel content validation is not one statistical task. A relevance rating,
an essential/not-essential judgment, and an item-objective congruence judgment
ask experts different questions and therefore support different indices.
`expert_validity()` uses an explicit `mode` so those designs are not treated as
interchangeable.

The three modes are:

- **relevance**: Aiken's V plus CVI and modified kappa;
- **essentiality**: Lawshe's CVR with exact binomial inference; and
- **congruence**: Rovinelli-Hambleton IOC.

All three are quantitative complements to qualitative expert comments, construct
coverage, comprehensibility review, and other parts of the content-validity
argument.

## Relevance: Aiken V, score intervals, and CVI

Suppose six experts rate item relevance from 1 (not relevant) to 4 (highly
relevant):

```{r relevance}
R <- matrix(
  c(4,4,4,4,4,4,
    4,4,4,3,4,4,
    4,3,4,4,3,4,
    3,3,4,3,2,3),
  nrow = 6,
  dimnames = list(NULL, paste0("Item", 1:4))
)

fit <- expert_validity(R, mode = "relevance", lo = 1, hi = 4)
fit
summary(fit)
```

Aiken's V rescales the bounded expert ratings to the 0-1 interval. The default
confidence interval is the score interval proposed by Penfield and Giacobbi
(2004), rather than a simulation-dependent bootstrap interval. Bootstrap
intervals remain available through the low-level function:

```{r aiken-bootstrap}
aikens_v(R, lo = 1, hi = 4, ci = "bootstrap", B = 200, seed = 1)
```

For CVI, the workflow dichotomizes ratings at `relevance_cut`. On a 1-4 scale
the default is 3, so ratings of 3 or 4 count as relevant. Declare a different
threshold if the study protocol used one.

The workflow reports common panel-size I-CVI guidelines (1.00 for panels of
3-5 experts and .78 for 6 or more) as **review aids**. They are not presented as
universal proof that an item is or is not content valid. Modified kappa provides
a chance-corrected complement to I-CVI.

At the scale level, S-CVI/Ave and S-CVI/UA are reported together. S-CVI/Ave is
generally less brittle than universal agreement, but both should be interpreted
alongside the distribution of item-level evidence.

## Essentiality: Lawshe CVR with exact critical values

Lawshe's task asks experts whether an item is essential. With twelve experts:

```{r essentiality}
expert_validity(c(10, 8, 6), mode = "essentiality", N = 12)
```

`cvr()` derives the smallest essential count whose one-sided binomial upper-tail
probability is no greater than `alpha`. This makes the panel-size dependency
explicit and follows the exact-probability logic revisited by Ayre and Scally
(2014).

Judge-by-item binary data can be supplied directly:

```{r essential-matrix}
E <- cbind(
  Item1 = c(1,1,1,1,1,1,1,1),
  Item2 = c(1,1,1,1,1,0,0,0)
)
expert_validity(E, mode = "essentiality")
```

A failure to clear the exact criterion is labeled `Review`, not automatic
deletion. Expert rationales and domain coverage matter when deciding whether an
item should be rewritten, retained for breadth, or removed.

## Congruence: item-objective alignment

IOC uses expert ratings of -1, 0, and +1 for item-objective congruence. A target
mapping lets the workflow compare intended and competing objectives:

```{r congruence}
d <- expand.grid(
  item = c("I1", "I2"),
  judge = 1:4,
  objective = c("A", "B")
)
d$target_objective <- ifelse(d$item == "I1", "A", "B")
d$score <- ifelse(d$objective == d$target_objective, 1, -1)

expert_validity(d, mode = "congruence")
```

The workflow reports target IOC, the strongest competitor, and their margin.
This is a diagnostic comparison, not a manufactured significance test. If no
target mapping is provided, all IOC cells are returned descriptively.

## Missing ratings

Missing data are never silently ignored by default. Set `na.rm = TRUE` only when
itemwise/cellwise deletion matches the study protocol. Effective expert counts
and missing counts are then reported so downstream interpretation uses the
actual panel size.

## Plotting expert evidence

Each mode uses a plot matched to the expert task rather than forcing unlike
indices into one generic chart.

```{r expert-plots, fig.width=7, fig.height=4}
plot(expert_validity(R, mode = "relevance", lo = 1, hi = 4))
plot(expert_validity(c(10, 8, 6), mode = "essentiality", N = 12))
plot(expert_validity(d, mode = "congruence"))
```

Relevance mode displays Aiken's V with its score interval and overlays I-CVI as
a separate marker. Essentiality mode displays observed CVR against the exact
panel-specific critical CVR. Congruence mode connects target IOC to the strongest
competitor so the alignment margin is visually explicit. These displays are
diagnostic summaries; they do not create new validity thresholds.

## Reporting

A concise methods/results description should identify:

1. who the experts were and why they were qualified;
2. the exact task and response scale;
3. the index and inference/CI procedure used;
4. the panel size, including item-specific missingness;
5. quantitative item and scale evidence; and
6. how expert comments, construct coverage, and comprehensibility informed the
   final item decisions.

A content-validity coefficient is evidence about a defined expert task. It is
not, by itself, a complete validity argument.

## References

Aiken, L. R. (1980). Content validity and reliability of single items or
questionnaires. *Educational and Psychological Measurement, 40*, 955-959.

Ayre, C., & Scally, A. J. (2014). Critical values for Lawshe's content validity
ratio: Revisiting the original methods of calculation. *Measurement and
Evaluation in Counseling and Development, 47*(1), 79-86.
https://doi.org/10.1177/0748175613513808

Penfield, R. D., & Giacobbi, P. R., Jr. (2004). Applying a score confidence
interval to Aiken's item content-relevance index. *Measurement in Physical
Education and Exercise Science, 8*(4), 213-225.
https://doi.org/10.1207/S15327841MPEE0804_3

Polit, D. F., Beck, C. T., & Owen, S. V. (2007). Is the CVI an acceptable
indicator of content validity? Appraisal and recommendations. *Research in
Nursing & Health, 30*(4), 459-467. https://doi.org/10.1002/nur.20199

Rovinelli, R. J., & Hambleton, R. K. (1977). On the use of content specialists
in the assessment of criterion-referenced test item validity. *Dutch Journal of
Educational Research, 2*, 49-60.
