---
title: "Getting Started with contentvalidR"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Getting Started with contentvalidR}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>")
set.seed(1)
```

```{r setup}
library(contentvalidR)
```

**Overview**

This vignette shows the two main workflows—**sort-based** (Psa/Csv) and **rating-based** (ANOVA + Tukey)—and **classic indices** (CVR, IOC, Aiken’s V, CVI), plus diagnostic tools.

**Sort-based (Psa, Csv, binomial)**
```{r}
toy_sort <- data.frame(
  item = rep(paste0("I", 1:4), each = 12),
  rater = rep(1:12, 4),
  target_construct   = rep(c("A","A","B","B"), each = 12),
  assigned_construct = c(
    sample(c("A","B"), 12, TRUE, c(.80,.20)),
    sample(c("A","B"), 12, TRUE, c(.65,.35)),
    sample(c("A","B"), 12, TRUE, c(.70,.30)),
    sample(c("A","B"), 12, TRUE, c(.45,.55))
  )
)
psa <- compute_psa(toy_sort)
csv <- compute_csv(toy_sort)
csv$decision <- vapply(seq_len(nrow(csv)), function(i) {
  csv_binom_test(csv$n_target[i], csv$n[i])$decision
}, character(1))
psa; csv
```

**Interpretation**
- **Psa** = share assigning the intended construct.
- **Csv** = margin of wins: $ \frac{n_{target} - max_{other}}{N} $.
- Binomial test (H0: $ p ≤ .5 $) flags items with above-chance targeting.

**Rating-based (ANOVA + Tukey)**
```{r}
set.seed(2)
toy_ratings <- data.frame(
  item = rep(paste0("I",1:4), each = 30),
  rater = rep(1:30, times = 4),
  construct = rep(rep(LETTERS[1:3], each = 10), times = 4),
  rating = c(
    rnorm(10, 4.3, 0.4), rnorm(20, 2.6, 0.6),
    rnorm(10, 4.0, 0.5), rnorm(20, 2.7, 0.6),
    rnorm(10, 2.6, 0.6), rnorm(10, 4.1, 0.5), rnorm(10, 2.6, 0.6),
    rnorm(20, 2.6, 0.6), rnorm(10, 4.0, 0.5)
  )
)
target_map <- c(I1="A", I2="A", I3="B", I4="C")
anova_content(toy_ratings, target_map = target_map)
```

**Interpretation**
- ANOVA $ p $ tests any mean difference across constructs for an item.
- Tukey confirms **target mean > all others** (strict convergent/discriminant).

**Classic indices**
```{r}
R <- matrix(sample(1:5, 5*6, replace = TRUE), nrow = 5)
aikens_v(R, lo = 1, hi = 5)

cvr(essential = c(8,10,5), N = 12)

M <- matrix(sample(0:1, 6*5, replace = TRUE, prob = c(.3,.7)), nrow = 6)
cvi(M)

ioc_df <- data.frame(
  item = rep(paste0("I",1:2), each = 9),
  judge = rep(1:3, times = 6),
  objective = rep(rep(LETTERS[1:3], each = 3), times = 2),
  score = sample(c(-1,0,1), 18, replace = TRUE)
)
ioc(ioc_df)
```

**Diagnostics & reproducibility**
```{r}
truth <- c(TRUE, TRUE, TRUE, FALSE)  # pretend "kept" after CFA
signal_detection(csv$decision == "significant", truth)

csv2_sig <- sample(c(TRUE, FALSE), nrow(csv), replace = TRUE)
reproducibility_phi(csv$decision == "significant", csv2_sig)
```

**Power quick-checks**
```{r}
simulate_csv_power(N = 20, true_p = 0.65, reps = 200)
simulate_anova_power(n_raters = 20, mean_diff = 0.5, sd = 1, k_constructs = 3, reps = 100)
```
