---
title: "Getting Started with contentvalidR"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Getting Started with contentvalidR}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>")
set.seed(1)
```

```{r setup}
library(contentvalidR)
```

## Overview

contentvalidR analyzes the evidence that judges and experts provide about a
scale's content before anyone answers it. Each kind of study has its own
workflow function, and each workflow has a guide:

| The study | Workflow | Guide |
|---|---|---|
| Judges sort each item into the construct it best represents | `sort_validity()` | `vignette("item-sort-validity")` |
| Judges rate each item against every construct definition | `rating_validity()` | `vignette("construct-rating-validity")` |
| Experts rate each item's relevance, essentiality, or congruence with an objective | `expert_validity()` | `vignette("expert-panel-validity")` |
| A panel rates the items over several Delphi rounds | `delphi_validity()` | `vignette("delphi-rounds")` |
| Do the conclusions depend on the particular judges? | `judge_validity()` | "Reading judge heterogeneity" in `vignette("reading-output")` |
| Do the items cover every cell of the blueprint? | `domain_validity()` | "Reading domain coverage" in `vignette("reading-output")` |

Two functions work across the workflows. `content_evidence()` sets several
review stages side by side, item by item, and draws them as figures
(`vignette("reporting-examples")`). `content_handoff()` carries the items a
review kept, with the evidence behind each decision, into the analysis of
response data (`vignette("handoff-to-empirical-validation")`, and
`vignette("one-item-set-both-stages")` for one item set followed through both
stages).

This vignette runs the first three workflows on small examples and points to
the rest. `vignette("reading-output")` explains every part of a printout, and
`vignette("design-and-reporting")` covers how many judges to recruit and what
to report.

## A common workflow contract

All six workflow objects expose `results`, `scale_summary`, `settings`,
`design`, and `details`. Their `results` tables also contain a common `status`
field: `Supported`, `Review`, `Insufficient data`, or `Descriptive only`. The
method's own decision word, such as `Retain` in an item sort, is kept beside it
as `recommendation`. This makes it possible to write reusable code across
workflows without pretending that a Howard-Melloy retention decision, a
construct-rating screening result, and an expert-panel judgment are
substantively identical.

## Item sort: Psa, Csv, and the exact test

Twelve judges sorted four items into three constructs, A, B, and C. Items I1
and I2 were written for A, and I3 and I4 for B:

```{r sort}
toy_sort <- data.frame(
  item = rep(paste0("I", 1:4), each = 12),
  rater = rep(1:12, 4),
  target_construct = rep(c("A", "A", "B", "B"), each = 12),
  assigned_construct = c(
    rep("A", 11), "C",
    rep("A", 10), "B", "C",
    rep("B", 7), rep("A", 4), "C",
    rep("B", 4), rep("A", 6), rep("C", 2)
  )
)
sort_fit <- sort_validity(toy_sort)
sort_fit
```

### Interpretation

- **Psa**, the proportion of substantive agreement, is the share of judges who
  assign the item to its intended construct: \(\text{Psa} = n_{target} / N\).
- **Csv**, the substantive-validity coefficient, is the intended construct's
  margin over its strongest competitor:
  \(\text{Csv} = (n_{target} - n_{competitor}) / N\). Here \(n_{target}\)
  counts the assignments to the intended construct, \(n_{competitor}\) those
  to the most-chosen other construct, and \(N\) the judges who sorted the item
  (Anderson & Gerbing, 1991). A negative Csv, as for I4, means a competing
  construct drew more judges than the intended one.
- The exact binomial test of Howard and Melloy (2016) asks whether the target
  count is too large for a target rate of .50 or less. With twelve judges an
  item needs ten target assignments, so I1 and I2 are retained and I3 and I4
  are flagged for review. The .50 is a benchmark rate, not the rate random
  sorting would give, which is 1 divided by the number of constructs.

The components are also available on their own:

```{r sort-components}
compute_csv(toy_sort)
csv_binom_test(n_c = 10, N = 12)
```

## Construct ratings: HTC, HTD, and the repeated-measures test

```{r rating}
set.seed(2)
toy_ratings <- expand.grid(
  item = c("I1", "I2", "I3"),
  rater = 1:16,
  construct = c("A", "B", "C")
)
toy_ratings$target_construct <- ifelse(toy_ratings$item == "I3", "B", "A")
toy_ratings$rating <- ifelse(
  toy_ratings$construct == toy_ratings$target_construct,
  pmin(5, pmax(1, round(rnorm(nrow(toy_ratings), 4.4, .6)))),
  pmin(5, pmax(1, round(rnorm(nrow(toy_ratings), 2.2, .7))))
)

rating_fit <- rating_validity(toy_ratings, scale_min = 1, scale_max = 5)
rating_fit
summary(rating_fit)
```

### Interpretation

- **HTC** summarizes definitional correspondence: how highly judges rated the
  item against its intended construct (Hinkin & Tracey, 1999; Colquitt et al.,
  2019).
- **HTD** summarizes definitional distinctiveness: how far the intended
  construct's ratings lead the orbiting constructs' ratings, on average.
- The repeated-measures ANOVA tests whether the construct-definition ratings
  differ for an item.
- Planned paired contrasts ask the direct screening question: is the target
  rating significantly higher than every orbiting rating?
- Scale-level HTC and HTD averages can be read against the empirical norms of
  Colquitt et al. (2019) when the judges are like the naive judges those norms
  came from.

## Expert panels: relevance, essentiality, and congruence

```{r expert}
expert_ratings <- matrix(
  c(4,4,4,4,4,4,
    4,4,4,3,4,4,
    4,3,4,4,3,4),
  nrow = 6,
  dimnames = list(NULL, paste0("Item", 1:3))
)
expert_fit <- expert_validity(expert_ratings, mode = "relevance", lo = 1, hi = 4)
expert_fit
summary(expert_fit)
```

Relevance, essentiality, and congruence are intentionally separate expert
tasks. Use `mode = "relevance"` for Aiken's V with the I-CVI and modified
kappa, `mode = "essentiality"` for Lawshe's CVR, and `mode = "congruence"` for
the index of item-objective congruence (IOC). `vignette("expert-panel-validity")`
walks through all three.

## Delphi rounds

When the same panel rates the items over several rounds, with anonymous
feedback in between, `delphi_validity()` answers two questions side by side:
whether enough experts agree about an item now (consensus, against a threshold
you fix before the study), and whether experts are still changing their
ratings (stability). It never folds one into the other. See
`vignette("delphi-rounds")`.

## Judges and domain coverage

Two workflows ask questions that no item-level index can reach.
`judge_validity()` returns one row per judge and asks whether the conclusions
depend on the particular judges who served. `domain_validity()` returns one row
per cell of the content blueprint and asks whether the items cover every cell,
including a cell no item was written for. Both are read in
`vignette("reading-output")`.

## From review to response data

`content_evidence()` brings several review stages together, such as a
relevance panel followed by an item sort, and draws a flow diagram and an
evidence profile for a paper or poster (`vignette("reporting-examples")`).
`content_handoff()` packages a review's decision for the empirical stage: the
items carried forward, the items held back and why, and two facts only you
know, the reverse-worded items and the response scale:

```{r handoff}
handoff <- content_handoff(sort_fit, reverse_keyed = character(0),
                           response_scale = c(1, 5))
handoff$items
```

`vignette("handoff-to-empirical-validation")` describes what the handoff
holds, and `vignette("one-item-set-both-stages")` follows one item set through
content review and the analysis of responses.

## Bundled reproducible examples

The package also installs deterministic CSV examples for the three workflow
families and all expert-panel modes. They are synthetic, contain no participant
data, and are regenerated from `data-raw/build-example-data.R` in the source
repository.

```{r bundled-data}
example_files <- c(
  "sort_example.csv",
  "rating_example.csv",
  "expert_relevance_example.csv",
  "expert_essentiality_example.csv",
  "expert_congruence_example.csv"
)
paths <- vapply(example_files, function(x) {
  system.file("extdata", x, package = "contentvalidR")
}, character(1))
file.exists(paths)
```

See `vignette("reporting-examples", package = "contentvalidR")` for
manuscript-ready reporting scaffolds built from those same files.

## Classic indices

```{r classic}
R <- matrix(sample(1:5, 5*6, replace = TRUE), nrow = 5)
aikens_v(R, lo = 1, hi = 5)

cvr(essential = c(8,10,5), N = 12)

M <- matrix(sample(0:1, 6*5, replace = TRUE, prob = c(.3,.7)), nrow = 6)
cvi(M)

ioc_df <- data.frame(
  item = rep(paste0("I",1:2), each = 9),
  judge = rep(1:3, times = 6),
  objective = rep(rep(LETTERS[1:3], each = 3), times = 2),
  score = sample(c(-1,0,1), 18, replace = TRUE)
)
ioc(ioc_df)
```

## Diagnostics and reproducibility

Two auxiliary helpers compare the sort's decisions with a later outcome. Here
the later outcome is invented: suppose a confirmatory factor analysis kept I1,
I2, and I3, and a second, independent sort supported I1, I2, and I4.

```{r diagnostics}
pretest_supported <- sort_fit$results$status == "Supported"
later_retained <- c(TRUE, TRUE, TRUE, FALSE)
signal_detection(pretest_supported, later_retained)

replication_supported <- c(TRUE, TRUE, FALSE, TRUE)
reproducibility_phi(pretest_supported, replication_supported)
```

Four items are far too few for either coefficient to mean much; the example
shows the calls. With so few items every expected count is below 5, so both
helpers report Fisher's exact *p*.

## Power quick-checks

`sort_power()` gives the exact probability that an item reaches the retention
count, for each number of judges and each rate at which you expect judges to
choose the target. `expert_power()` does the same for the I-CVI criterion of an
expert panel; see `vignette("design-and-reporting")`.

```{r power}
sort_power(N = c(20, 30), true_p = c(.65, .75))
```

## References

Anderson, J. C., & Gerbing, D. W. (1991). Predicting the performance of
measures in a confirmatory factor analysis with a pretest assessment of their
substantive validities. *Journal of Applied Psychology, 76*(5), 732–740.
https://doi.org/10.1037/0021-9010.76.5.732

Colquitt, J. A., Sabey, T. B., Rodell, J. B., & Hill, E. T. (2019). Content
validation guidelines: Evaluation criteria for definitional correspondence and
definitional distinctiveness. *Journal of Applied Psychology, 104*(10),
1243–1265. https://doi.org/10.1037/apl0000406

Hinkin, T. R., & Tracey, J. B. (1999). An analysis of variance approach to
content validation. *Organizational Research Methods, 2*(2), 175–186.
https://doi.org/10.1177/109442819922004

Howard, M. C., & Melloy, R. C. (2016). Evaluating item-sort task methods: The
presentation of a new statistical significance formula and methodological best
practices. *Journal of Business and Psychology, 31*(1), 173–186.
https://doi.org/10.1007/s10869-015-9404-y
