---
title: "Getting Started with contentvalidR"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Getting Started with contentvalidR}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>")
set.seed(1)
```

```{r setup}
library(contentvalidR)
```

**Overview**

This vignette introduces the package's two primary content-pretesting workflows—**item sorting** and **construct ratings**—plus complementary expert-panel indices. `sort_validity()` and `rating_validity()` both organize item-level screening evidence and target-scale normative summaries while keeping substantive decisions separate from statistical flags.

**Sort-based (Psa, Csv, binomial)**
```{r}
toy_sort <- data.frame(
  item = rep(paste0("I", 1:4), each = 12),
  rater = rep(1:12, 4),
  target_construct   = rep(c("A","A","B","B"), each = 12),
  assigned_construct = c(
    sample(c("A","B"), 12, TRUE, c(.80,.20)),
    sample(c("A","B"), 12, TRUE, c(.65,.35)),
    sample(c("A","B"), 12, TRUE, c(.70,.30)),
    sample(c("A","B"), 12, TRUE, c(.45,.55))
  )
)
psa <- compute_psa(toy_sort)
csv <- compute_csv(toy_sort)
csv$decision <- vapply(seq_len(nrow(csv)), function(i) {
  csv_binom_test(csv$n_target[i], csv$n[i])$decision
}, character(1))
psa; csv
```

**Interpretation**
- **Psa** = share assigning the intended construct.
- **Csv** = margin of wins: $ \frac{n_{target} - max_{other}}{N} $.
- Binomial test (H0: $ p ≤ .5 $) flags items with above-chance targeting.

**Construct-rating workflow (HTC, HTD, repeated-measures ANOVA)**
```{r}
set.seed(2)
toy_ratings <- expand.grid(
  item = c("I1", "I2", "I3"),
  rater = 1:16,
  construct = c("A", "B", "C")
)
toy_ratings$target_construct <- ifelse(toy_ratings$item == "I3", "B", "A")
toy_ratings$rating <- ifelse(
  toy_ratings$construct == toy_ratings$target_construct,
  pmin(5, pmax(1, round(rnorm(nrow(toy_ratings), 4.4, .6)))),
  pmin(5, pmax(1, round(rnorm(nrow(toy_ratings), 2.2, .7))))
)

rating_fit <- rating_validity(toy_ratings, scale_min = 1, scale_max = 5)
rating_fit
summary(rating_fit)
```

**Interpretation**
- **HTC** summarizes definitional correspondence with the intended construct.
- **HTD** summarizes distinctiveness from orbiting constructs.
- The repeated-measures ANOVA tests whether construct-definition ratings differ for an item.
- Planned paired contrasts ask the direct screening question: is the target rating significantly higher than every orbiting rating?
- Scale-level HTC/HTD averages can be interpreted using Colquitt et al. (2019) empirical norms when the judge population matches their intended use.

**Classic indices**
```{r}
R <- matrix(sample(1:5, 5*6, replace = TRUE), nrow = 5)
aikens_v(R, lo = 1, hi = 5)

cvr(essential = c(8,10,5), N = 12)

M <- matrix(sample(0:1, 6*5, replace = TRUE, prob = c(.3,.7)), nrow = 6)
cvi(M)

ioc_df <- data.frame(
  item = rep(paste0("I",1:2), each = 9),
  judge = rep(1:3, times = 6),
  objective = rep(rep(LETTERS[1:3], each = 3), times = 2),
  score = sample(c(-1,0,1), 18, replace = TRUE)
)
ioc(ioc_df)
```

**Diagnostics & reproducibility**
```{r}
truth <- c(TRUE, TRUE, TRUE, FALSE)  # pretend "kept" after CFA
signal_detection(csv$decision == "significant", truth)

csv2_sig <- sample(c(TRUE, FALSE), nrow(csv), replace = TRUE)
reproducibility_phi(csv$decision == "significant", csv2_sig)
```

**Power quick-checks**
```{r}
sort_power(N = c(20, 30), true_p = c(.65, .75))
```
