---
title: "Manuscript-Ready Reporting Examples"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Manuscript-Ready Reporting Examples}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>")
library(contentvalidR)
read_example <- function(name) {
  utils::read.csv(
    system.file("extdata", name, package = "contentvalidR"),
    stringsAsFactors = FALSE
  )
}
```

## Purpose

This vignette provides reporting scaffolds for the three flagship workflows. The
examples are intentionally conservative: statistical screening is described as
**evidence for retention or review**, not as proof that an item or scale is
content valid. Final decisions should also consider construct-domain coverage,
item wording, qualitative judge feedback, and the intended use of the measure.

The bundled CSV files are synthetic, deterministic, and regenerated from
`data-raw/build-example-data.R`. They are useful for reproducing the examples
and for seeing the expected input shape before analyzing a new study.

## Item-sort study

```{r sort-fit}
sort_dat <- read_example("sort_example.csv")
sort_fit <- sort_validity(sort_dat)
sort_sum <- summary(sort_fit)
sort_fit$results
sort_fit$scale_summary
```

### Methods scaffold

Report who completed the sort, how the construct definitions were presented,
the available assignment alternatives, and the a priori screening rule. A
concise methods statement can follow this structure:

> Candidate items were evaluated in an item-sort pretest in which judges
> assigned each item to the construct definition that best represented its
> content. We quantified definitional correspondence using the proportion of
> substantive agreement (Psa) and definitional distinctiveness using the
> substantive-validity coefficient (Csv). Item-level screening used the exact
> target-assignment test described by Howard and Melloy (2016), with the null
> target-assignment probability and alpha specified a priori. Target-scale mean
> Psa and Csv were interpreted against Colquitt et al. (2019) norms only when the
> judge population matched the intended use of those norms.

### Results scaffold

For this bundled example, `r sort_fit$design$n_raters` judges evaluated
`r sort_fit$design$n_items` items. `r sort_sum$n_supported` items met the exact
screening criterion, `r sort_sum$n_review` were flagged for review, and
`r sort_sum$n_insufficient` had insufficient usable assignments.

A manuscript table can usually be built directly from:

```{r sort-table}
sort_fit$results[c(
  "item", "target", "n", "n_target", "competitor",
  "psa", "csv", "p_value", "status", "recommendation"
)]
```

Do not report `Review` as synonymous with deletion. A review flag identifies an
item for substantive inspection; retaining an item for domain coverage can be a
reasonable decision when that rationale is documented.

## Construct-rating study

```{r rating-fit}
rating_dat <- read_example("rating_example.csv")
rating_fit <- rating_validity(rating_dat, scale_min = 1, scale_max = 5)
rating_sum <- summary(rating_fit)
rating_fit$results
rating_fit$scale_summary
```

### Methods scaffold

> Judges rated every candidate item against each focal and orbiting construct
> definition using the same response scale. We summarized correspondence with
> HTC and distinctiveness with HTD. Because the same judges rated the competing
> definitions, item-level inference used a repeated-measures design. Planned
> paired contrasts compared each item's intended definition with every orbiting
> definition; omnibus Greenhouse-Geisser-corrected inference was used when
> applicable. Scale-level HTC/HTD norms from Colquitt et al. (2019) were treated
> as empirical benchmarks rather than universal item cutoffs.

### Results scaffold

The example contains `r rating_fit$design$n_items` items rated by
`r rating_fit$design$n_raters` judges against
`r rating_fit$design$n_constructs_observed` construct definitions.
`r rating_sum$n_supported` items were supported by the complete screening rule
and `r rating_sum$n_review` were flagged for review.

```{r rating-table}
rating_fit$results[c(
  "item", "target", "n_complete", "strongest_competitor",
  "htc", "htd", "p_value", "max_contrast_p", "status", "recommendation"
)]
```

For review items, report the strongest orbiting competitor. That information
turns a generic statement about weak distinctiveness into a specific diagnostic
about where construct overlap may be occurring.

## Expert-panel study

### Relevance

```{r expert-relevance}
expert_rel <- read_example("expert_relevance_example.csv")
expert_rel_matrix <- as.matrix(expert_rel[setdiff(names(expert_rel), "expert")])
expert_fit <- expert_validity(
  expert_rel_matrix,
  mode = "relevance",
  lo = 1,
  hi = 4
)
expert_sum <- summary(expert_fit)
expert_fit$results
expert_fit$scale_summary
```

> Experts rated the relevance of each candidate item on a bounded ordinal
> scale. We summarized relevance using Aiken's V with Penfield-Giacobbi score
> confidence intervals and calculated I-CVI with Polit-Beck-Owen modified kappa.
> S-CVI/Ave and S-CVI/UA were reported at the scale level. Panel-size CVI
> guidelines were used as review aids and were considered alongside written
> expert feedback and construct coverage.

In this example, `r expert_fit$design$n_judges` experts evaluated
`r expert_fit$design$n_items` items. The workflow identifies
`r expert_sum$n_supported` supported items and `r expert_sum$n_review` review
items under its quantitative rules.

### Essentiality

```{r expert-essentiality}
expert_ess <- read_example("expert_essentiality_example.csv")
expert_ess_matrix <- as.matrix(expert_ess[setdiff(names(expert_ess), "expert")])
ess_fit <- expert_validity(expert_ess_matrix, mode = "essentiality")
ess_fit$results
```

For an essentiality task, state that CVR is tied to a different expert judgment
than relevance. Report the effective panel size and exact critical essential
count for each item rather than borrowing a relevance/CVI threshold.

### Congruence

```{r expert-congruence}
expert_ioc <- read_example("expert_congruence_example.csv")
ioc_fit <- expert_validity(expert_ioc, mode = "congruence")
ioc_fit$results
```

For IOC, report the intended objective, target IOC, strongest competing
objective, and target-minus-competitor margin. The margin is diagnostic evidence
about alignment; it is not a newly invented significance test.

## Minimum reproducibility statement

At minimum, a manuscript or supplement should identify the package version and
record the analysis settings that determine the results. For a fitted workflow:

```{r reproducibility}
packageVersion("contentvalidR")
sort_fit$settings
sort_fit$design
```

A strong reproducibility supplement should also archive the item wording,
construct definitions, judge instructions, anonymized response data when
permitted, and the script that reproduces all tables and figures.

## Language to avoid

Avoid claims such as "the scale was proven content valid" or "items failing the
cutoff were invalid." The workflows quantify evidence from a defined pretest.
They do not replace a theory-based definition of the construct domain or the
researcher's responsibility to justify substantive item decisions.
